### 2021-01-07 install.Rmd
Initial commit

---
title: "Installation of Required Software"
author: "Aaron Peikert"
date: "05/01/2021"
output:
  github_document:
    number_sections: true
repro:
  packages:
    - here
  scripts:
    - R/link.R
---

```{r setup, include=FALSE}
source(here::here("R", "link.R"))
```

# Why you want to use package managers

Installing software can be a painful experience, especially if you have to sift through the documentation for multiple platforms.
A package manager helps you to install, update and remove software on your computer, automating large parts of the process for you.
Most importantly, the installation process is the same for every software, not unlike the appstore on your phone.
Unfortunately though, it is not as pleasantly designed, but is used through the terminal.
If the thought of using a terminal lets you hesitate, rest assured it is easier than you might think.

# Windows 10

## How to open a terminal on Windows 10?

1. Press Windows 10 key + X.
2. Click on "Windows Powershell" or "Windows Powershell (Admin)" (if you are using Chocolately)

## How to install a package manager for Windows 10?

`r link("https://chocolatey.org/", "Chocolately")` is a great package manager for Windows 10.
You should follow these installation instructions: `r link("https://chocolatey.org/docs/installation")`, but if you are in a hurry:

1. Press Windows key + X, click on "Windows Powershell (Admin)".
2. Paste: `Set-ExecutionPolicy Bypass -Scope Process -Force; [System.Net.ServicePointManager]::SecurityProtocol = [System.Net.ServicePointManager]::SecurityProtocol -bor 3072; iex ((New-Object System.Net.WebClient).DownloadString('https://chocolatey.org/install.ps1'))`
3. Press enter.
4. Close the PowerShell

## How to install R + Rstudio with a package manager?

Since you probably already have R and RStudio you may skip this step.

1. Open a new terminal (Press Windows key + X, click on "Windows Powershell (Admin)").
2. Paste `choco install -y r r.studio`
3. Press enter

## How to install Git, Make, and Docker with a package manager?

Please note that Docker does not work on Windows 10 Home, you need Windows Pro or Education.

1. Open a new terminal (Press Windows key + X, click on "Windows Powershell (Admin)").
2. Paste `choco install -y git make docker`
3. Press enter.

# Mac OS

## How to open a terminal on Mac OS?

1. Press Command + Space.
2. Type "Terminal".
3. Press enter.

## How to install a package manager for Mac OS?

`r link("https://brew.sh", "Homebrew")` is a great package manager for Mac OS. To install it:

1. Open a new terminal (Press Command + Space, type "Terminal", press enter).
2. Paste `xcode-select --install` and press enter to install xcode.
3. Paste: `/bin/bash -c "$(curl -fsSL https://raw.githubusercontent.com/Homebrew/install/HEAD/install.sh)"`.
4. Press enter.
5. Close the terminal.

## How to install R + Rstudio with a package manager?

Since you probably already have R and RStudio you may skip this step.

1. Open a new terminal (Press Command + Space, type "Terminal", press enter).
2. Paste `brew install r` and press enter to install R.
3. Paste `brew install --cask rstudio` and press enter to install RStudio.

## How to install Git, Make, and Docker with a package manager?

1. Open a new terminal (Press Command + Space, type "Terminal", press enter).
2. Paste `brew install git make docker`.
3. Press enter.

```{r link_index, results='asis', echo=FALSE}
link_index()
```