### 2021-05-04 README.Rmd
add reproduction statement to readme closes #27

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# A Hitchhiker's Guide to Reproducible Research
<!-- badges: start -->
[![Project Status: Concept – Minimal or no implementation has been done yet, or the repository is only intended to be a limited example, demo, or proof-of-concept.](https://www.repostatus.org/badges/latest/concept.svg)](https://www.repostatus.org/#concept)
<!-- badges: end -->

WIP

## How to reproduce this manuscript

To reproduce this project Git, Make, and Docker is required (see [the installation guide](https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md)).

Open the terminal, download the repository, and enter the directory:

```bash
git clone https://github.com/aaronpeikert/repro-tutorial.git
cd repro-tutorial
```

Then build the Docker image, and run Make:

```bash
make docker &&
make -B DOCKER=TRUE 
```

## Code of Conduct

Please note that the repro-tutorial project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html).
By contributing to this project, you agree to abide by its terms.