### 2021-06-08 abstract.Rmd
include originator as beneficiary

Computational reproducibility is the ability to obtain identical results from the *same* data with the *same* computer code. It is a building block for transparent and cumulative science because it enables the originator and other researchers, on other computers and later in time, to reproduce and thus understand how results came about while, at the same time, it avoids a variety of errors that may lead to erroneous reporting of statistical and computational results. In this tutorial, we demonstrate how the R package `repro` supports researchers in creating fully computationally reproducible research projects with a focus on current Open Science practices; starting from sketching the research proposal, over preregistering an analysis as executable code with simulated data, until the creation and dissemination of the final scientific report.