### 2021-07-20 preregistration.Rmd
dont describe example, directly include it

---
title: "A Preregistered Comparison of Male and Female Machiavellianism"
date: "7/10/2021"
author:
  - name: Aaron Peikert
    affil: 1, *
    orcid: 0000-0001-7813-818X
  - name: Caspar J. Van Lissa
    affil: 2, 3
    orcid: 0000-0002-0808-5024
  - name: Andreas M. Brandmaier
    affil: 1, 4
    orcid: 0000-0001-8765-6982
affiliation:
  - num: 1
    address: |
      Center for Lifespan Psychology---Max Planck Institute for Human Development
      Lentzeallee 94, 14195 Berlin, Germany
  - num: 2
    address: |
      Department of Methodology & Statistics---Utrecht University faculty of Social and Behavioral Sciences, Utrecht, Netherlands
  - num: 3
    address: |
      Open Science Community Utrecht, Utrecht, Netherlands
  - num: 4
    address: |
      Max Planck UCL Centre for Computational Psychiatry and Ageing Research
      Berlin, Germany and London, UK
# firstnote to eighthnote
correspondence: |
  peikert@mpib-berlin.mpg.de
journal: psych
type: tutorial
status: submit
bibliography: preregistration.bib
abstract: |
  We test for group differences in machiavellism between self identified males and females in the MACH-IV in a well powered preregistered study.
output:
  bookdown::markdown_document2:
    base_format: rticles::mdpi_article
header-includes:
  -   \def\baselinestretch{0.98}
---

\setlength{\bibsep}{0cm}
\titlespacing{\section}{.1cm}{.1cm}{.1cm}

```{r setup, include=FALSE}
library(tidyverse)
library(report)
source(here::here("R", "simulation_funs.R"))
```

<!--begin-theory-->
# Theoretical Background

There is extensive literature reporting differences in the dark triad (narcissism, machiavellianism, and psychopathy) between self identified males and females.
@muris2017 note that studies focusing on gender differences in machiavellianism are sparse.
We therefore want to add to this body of literature.
<!--end-theory-->

```{r, echo=FALSE}
simulation_results <-
  readr::read_csv(here::here("data", "simulation_results.csv"), col_types = "ddddddd")
```

<!--begin-method-->
# Method

We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study [@simmons2012]. We use data available from [openpsychometrics.org](https://openpsychometrics.org/_rawdata/) from the online version of the MACH-IV[@christie1970] and included participants that have responded to at least one machiavellianism item and reported their gender as either "male" or "female".
<!--end-method-->

<!--begin-ttest-->
We conduct a Student's t-test [@studentProbableErrorMean1908] with Welch's correction [@welchGeneralizationStudentProblem1947] an the average of machiavellism items between the binary-coded gender groups.
If the skew of this average is greater than 1.0 we conduct a supposedly more robust Mann--Whitney--Wilcoxon test [@Wilcoxon1945] instead.
<!--end-ttest-->
<!--begin-power-code-->
```{r}
choosen_power <- .8
choosen_d <- .2
minn <-
  filter(filter(simulation_results, d == choosen_d, power > choosen_power),
         n == min(n))$n
```
<!--end-power-code-->
<!--begin-power-->
A simulation we conducted indicated that with a sample size of
`r minn` for an alpha of .05 (two-sided) we archive at least `r choosen_power*100`% power assuming an effect size of `r choosen_d`.
<!--end-power-->

<!--begin-results-->
# Results

```{r, echo=FALSE, results='asis'}
real_data <- here::here("data", "data.csv")
simulated <- !fs::file_exists(real_data)
if(simulated){
  cat("\\textcolor{red}{This document and the results presented herein are based
      on simulated data and should not be interpreted.}")
  set.seed(1234)
  mach <- simulate_data(900, 8, 0.3, 10)
} else {
  mach <- readr::read_csv(real_data)
}
```
<!--end-results-->

<!--begin-report-->
```{r, echo=FALSE, results='asis'}
report_analysis(planned_analysis(mach))
```
<!--end-report-->

# Discussion

\textcolor{red}{This document was created before we have gathered any data, we therefore have nothing to discuss.}

\newpage
# References