### 2021-07-30 abstract.Rmd
add PAC to abstract closes #131

Computational reproducibility is the ability to obtain identical results from the *same* data with the *same* computer code. It is a building block for transparent and cumulative science because it enables the originator and other researchers, on other computers and later in time, to reproduce and thus understand how results came about while, at the same time, it avoids a variety of errors that may lead to erroneous reporting of statistical and computational results. In this tutorial, we demonstrate how the R package `repro` supports researchers in creating fully computationally reproducible research projects with tools from the software engineering community. Building upon this notion of fully automated reproducibility we present several application including the preregistration of research plans with code (Preregistration as Code, PAC). PAC eschews all ambiguity of traditional preregistration and offers several other advantages.