### 2021-08-09 preregistration.Rmd
Necessary deviations from the preregistration

---
title: "Preregistered as Code: Gender differences in Machiavellianism"
date: "7/10/2021"
author:
  - name: Aaron Peikert
    affil: 1, *
    orcid: 0000-0001-7813-818X
  - name: Caspar J. Van Lissa
    affil: 2, 3
    orcid: 0000-0002-0808-5024
  - name: Andreas M. Brandmaier
    affil: 1, 4
    orcid: 0000-0001-8765-6982
affiliation:
  - num: 1
    address: |
      Center for Lifespan Psychology---Max Planck Institute for Human Development
      Lentzeallee 94, 14195 Berlin, Germany
  - num: 2
    address: |
      Department of Methodology & Statistics---Utrecht University faculty of Social and Behavioral Sciences, Utrecht, Netherlands
  - num: 3
    address: |
      Open Science Community Utrecht, Utrecht, Netherlands
  - num: 4
    address: |
      Max Planck UCL Centre for Computational Psychiatry and Ageing Research
      Berlin, Germany and London, UK
# firstnote to eighthnote
correspondence: |
  peikert@mpib-berlin.mpg.de
journal: psych
type: tutorial
status: submit
bibliography: preregistration.bib
abstract: |
  We test for group differences in machiavellianism between self-identified males and females in the MACH-IV in a well powered preregistered study.
keywords: |
  preregistration as code, power simulation, machiavellianism, gender
output: rticles::mdpi_article
header-includes:
  - \def\baselinestretch{0.98}
  - \usepackage{draftwatermark}
---

\setlength{\bibsep}{0cm}
\titlespacing{\section}{.1cm}{.1cm}{.1cm}
\SetWatermarkText{PAC}

```{r setup, include=FALSE}
library(tidyverse)
library(report)
source(here::here("R", "simulation_funs.R"))
```

<!--begin-theory-->
# Theoretical Background

Machiavellianism describes a personality dimension characterized by a cynical disregard of morals in the pursuit of one's own interest, e.g. through manipulation [@christie1970].
There is extensive literature reporting differences in the dark triad (narcissism, machiavellianism, and psychopathy) between self-identified males and females [@muris2017] but only few studies focus solely on machiavellianism.
We aim to replicate the finding that males tend to have higher machiavellianism scores [@muris2017].
<!--end-theory-->
This research question serves as a testbed for preregistration as code (PAC) to examine if PAC is feasible under realistic conditions.
In the PAC paradigm, all analysis code is written before real data are gathered or accessed the first time.

```{r, echo=FALSE}
simulation_results <-
  readr::read_csv(here::here("data", "simulation_results.csv"), col_types = "ddddddd")
```

<!--begin-method-->
# Method

We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study [cf. @simmons2012]. We use data available from [openpsychometrics.org](https://openpsychometrics.org/_rawdata/) from the online version of the MACH-IV[@christie1970] and included participants that have responded to at least one machiavellianism item and reported their gender as either "male" or "female".
<!--end-method-->

<!--begin-ttest-->
We conduct a Student's t-test [@studentProbableErrorMean1908] with Welch's correction [@welchGeneralizationStudentProblem1947] of the average of machiavellianism items between the binary-coded gender groups.
If the skew of this average is greater than 1.0 we conduct a supposedly more robust Mann--Whitney--Wilcoxon test [@Wilcoxon1945] instead.
<!--end-ttest-->
```{r, echo=FALSE}
choosen_power <- .8
choosen_d <- .2
minn <-
  filter(filter(simulation_results, d == choosen_d, power > choosen_power),
         n == min(n))$n
```
<!--begin-power-->
A simulation we conducted indicated that with a sample size of
`r minn` for an alpha of .05 (two-sided) we achieve at least `r choosen_power*100`% power assuming a standardized effect size of d=`r choosen_d`.
<!--end-power-->

<!--begin-results-->
# Results

```{r, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
real_data <- here::here("data", "data.csv")
simulated <- !fs::file_exists(real_data)
if(simulated){
  cat("\\textcolor{red}{The results are based on simulated data and must not be
      interpreted. They only serve to illustrate the result of the preregistered
      code.}")
  set.seed(1234)
  mach <- simulate_data(900, 8, 0.3, 10)
} else {
  mach <- readr::read_delim(real_data, delim = "\t")
  # only keep MACH items + gender
  mach <- dplyr::select(mach, dplyr::matches("^Q\\d+A$"), gender)
  # code gender according to codebook (3 would be other)
  mach <-
    dplyr::mutate(mach, gender = factor(
      gender,
      levels = 1:2,
      labels = c("male", "female")
    ))
}
```
<!--end-results-->

<!--begin-report-->
```{r, echo=FALSE, results='asis'}
report_analysis(planned_analysis(mach))
```
<!--end-report-->

# Discussion

We created this document to illustrate Preregistration as Code.
We therefore do not discuss the results.

\newpage
# References