### 2021-08-19 README.Rmd
add altmetric

---
output: github_document
repro:
  images:
    - images/nutshell.svg
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# A Hitchhiker's Guide to Reproducible Research
<!-- badges: start -->
<script type='text/javascript' src='https://d1bxh8uas1mnw7.cloudfront.net/assets/embed.js'></script>

<table>
<tbody>
  <tr>
    <td><div data-badge-popover="right" data-badge-type="1" data-doi="10.31234/osf.io/fwxs4" data-hide-no-mentions="true" class="altmetric-embed"></div></td>
    <td>Preregistration as Code: <a href="https://doi.org/10.5281/zenodo.5170740" target="_blank" rel="noopener noreferrer">[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.5170740.svg)](https://doi.org/10.5281/zenodo.5170740)</a></td>
  </tr>
  <tr>
    <td><a href="https://www.repostatus.org/#active" target="_blank" rel="noopener noreferrer">[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)</a></td>
    <td>Results after Preregistration: <a href="https://doi.org/10.5281/zenodo.5171678">[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.5171678.svg)](https://doi.org/10.5281/zenodo.5171678)</a></td>
  </tr>
</tbody>
</table>

<!-- badges: end -->

## How to reproduce this manuscript

To reproduce this project Git, Make, and Docker is required (see [the installation guide](https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md)).

Open the terminal, download the repository, and enter the directory:

```bash
git clone https://github.com/aaronpeikert/repro-tutorial.git
cd repro-tutorial
```

Then build the Docker image, and run Make:

```bash
make docker &&
make -B DOCKER=TRUE 
```

## Abstract

```{r abstract, child = 'abstract.Rmd'}
```

```{r, echo=FALSE}
knitr::include_graphics("images/nutshell.svg")
```

## Code of Conduct

Please note that the repro-tutorial project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html).
By contributing to this project, you agree to abide by its terms.