### 2021-09-23 manuscript.Rmd

---
title: "Reproducible Research in R: A tutorial on how to do the same thing more than once"
author:
  - name: Aaron Peikert
    affil: 1, *
    orcid: 0000-0001-7813-818X
  - name: Caspar J. Van Lissa
    affil: 2, 3
    orcid: 0000-0002-0808-5024
  - name: Andreas M. Brandmaier
    affil: 1, 4
    orcid: 0000-0001-8765-6982
affiliation:
  - num: 1
    address: |
      Center for Lifespan Psychology---Max Planck Institute for Human Development
      Lentzeallee 94, 14195 Berlin, Germany
  - num: 2
    address: |
      Department of Methodology & Statistics---Utrecht University faculty of Social and Behavioral Sciences, Utrecht, Netherlands
  - num: 3
    address: |
      Open Science Community Utrecht, Utrecht, Netherlands
  - num: 4
    address: |
      Max Planck UCL Centre for Computational Psychiatry and Ageing Research
      Berlin, Germany and London, UK
# firstnote to eighthnote
correspondence: |
  peikert@mpib-berlin.mpg.de
journal: psych
type: tutorial
status: submit
bibliography: temp.bib
simplesummary: |
  Reproducibility has long been considered integral to the scientific method.
  Something is called reproducible when an independent person obtains the same results from the same data.
  Until recently, detailed descriptions of methods and analyses were the primary instrument for ensuring scientific reproducibility.
  Technological advancements now enable scientists to achieve a more comprehensive standard;
  one in which any individual can be granted access to a digital research repository, and reproduce the analyses from the raw data to the final report including all relevant statistical analyses with a single command.
  This method has far-reaching implications for scientific archiving,
  reproducibility and replication,
  scientific productivity,
  and the credibility and reliability of scientific findings.
  One obstacle preventing the widespread adoption of this method is that the underlying technological advancements
  are complicated to use.
  This paper introduces `repro`, an R-package, which guides researchers in the installation and use of the tools required for making a research project reproducible.
  Finally, we suggest the use of the proposed tools for the preregistration of study plans as reproducible computer code (preregistration as code; PAC).
  Since computer code represents the planned analyses exactly as they will be executed, it is more precise than natural language descriptions of those analyses, which merely complement the PAC as a more readable summary. 
  PAC circumvents the shortcomings of ambiguous preregistrations that may give researchers undesired degrees of freedom.
  Hence, reproducibility made convenient with automation has a wide range of applications to accelerate scientific progress.
abstract: "`r tryCatch(trimws(readr::read_file(here::here('abstract.Rmd'))))`"
keywords: |
  open science; compuational reproducibility; preregistration; R; R Markdown; Make; GitHub; Docker
acknowledgement: |
  The authors received no financial support for the research, authorship, and/or publication of this article.
authorcontributions: |
  Aaron Peikert took the lead in writing and provided the initial draft of the manuscript.
  Andreas Brandmaier and Caspar J. Van Lissa contributed further ideas, critical feedback, and revisions of the original manuscript.
  Furthermore, we would like to thank Maximilian Stefan Ernst (not an author) for his contributions to the code for the simulation study.
conflictsofinterest: |
  The authors declare no conflict of interest.
abbreviations:
  - short: PAC
    long: Preregistration as Code
  - short: Gb
    long: Gigabyte
  - short: Kb
    long: Kilobyte
  - short: CRAN
    long: Comprehensive R Archive Network
  - short: WORCS
    long: Workflow for Open Reproducible Code in Science
repro:
  packages:
    - tidyverse
    - usethis
    - gert
    - aaronpeikert/repro@27e7d22
    - here
    - rstudio/webshot2@f62e743
    - targets
    - renv
    - slider
    - patchwork
    - knitr
    - pander
    - lavaan
    - furrr
    - future.batchtools
    - rticles
    - moments
    - report
  scripts:
    - R/simulation.R
    - R/simulation_funs.R
    - R/link.R
  data:
    - data/simulation_results.csv
  images:
    - images/nutshell.pdf
  files:
    - references.bib
output:
  bookdown::markdown_document2:
    base_format: rticles::mdpi_article
header-includes:
   - \DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}, numbers=left}
---

```{r setup, include=FALSE}
library(repro)
automate_load_packages()
automate_load_scripts()
source(here::here("R", "link.R"))
source(here::here("R", "simulation_funs.R"))
```

```{r echo=FALSE}
# define a print method for objects of the class data.frame
knit_print.data.frame = function(x, ...) {
  res = paste(c('', '', kable(x)), collapse = '\n')
  asis_output(res)
}
# register the method
registerS3method("knit_print", "data.frame", knit_print.data.frame)
```

# Introduction

Scientists increasingly strive to make research data, materials, and analysis code openly available.
Sharing these types of digital research products has the potential to increase scientific efficiency by enabling researchers to learn from each other and reuse materials, and to increase scientific reliability by facilitating the review and replication of published research results.
To some extent these potential benefits are contingent on whether these digital research products are reproducible. 
Reproducibility can be defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code [see @Peikert2019 for details].
The credibility of empirical research results hinges on their objectivity.
Objectivity in this context means "in principle it [the research finding] can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p. 22][^popper].
Only something reproducible can be tested and understood by anyone.
Therefore, reproducibility has long been considered integral to empirical research.
Unfortunately, despite increasing commitment to open science practices and good will, many projects can not yet be reproduced by other research teams [@obels2020].
This is because making a research project with all code, data, and materials reproducible comes with a variety of challenges (e.g., missing software or ambiguous documentation of all steps taken), and there is a dire need for reproducibility standards.
With the technological advancement, it is now possible to make all digital products related to a research project available so that a research project can be automatically reproduced with minimal effort.

We pursue two aims.
First, we want to introduce researchers to a notion of automated reproducibility that requires, apart from the setup of software, no manual steps.
Second, we discuss opportunities such automated reproducibility affords to improve not only how research outputs are created and recreated, but also how research itself is conducted.
With regard to the first goal, we present a tutorial that is aimed at researchers who intend to ensure the reproducibility of a research project and are willing to make relevant code, data, and materials available, whether publicly or on request.
In this tutorial, we introduce the reader to a workflow that maximizes the potential reproducibility of a research project.
The proposed workflow builds on several technical solutions originating from software engineering to automate reproduction. 
We discuss how these solutions address four common threats to reproducibility [see @Peikert2019 for details] and how they interact to form a coherent workflow.
We further demonstrate how the R package `repro` supports researchers in setting up the required software and in adopting this workflow.
In pursuit of the second aim we present a strictly reproducible and unambiguous form of preregistration [@NosekRevolution2018] that builds upon an implementation of this reproducible workflow, the so-called *preregistration as code* (PAC).
PAC involves preregistering the intended analysis code and the bulk of the manuscript as a dynamic document, including typical sections like Introduction, Methods, and Results.
The resulting dynamic document closely resembles the final manuscript, but temporarily uses simulated data to generate placeholder results (e.g. figures, tables and statistics).
With these placeholders, authors preregister the exact presentation of their results without having seen real data yet.
Once the empirical data are analyzed, the Results are then updated automatically, and the Discussion can be written to finalize the report.

[^popper]: In fact, @popperLogicScientificDiscovery2002's famous criterion of demarcation is built around "inter-subjective testing", a concept which he later generalized to inter-subjective criticism (Footnote *1, p. 22). Somewhat confusingly, @popperLogicScientificDiscovery2002 uses the term "Reproducibility" for what we would call "Replication".

Scientific organizations and funding bodies increasingly demand transparent sharing of digital research products, and researchers are increasingly willing to do so.
However, although the sharing of such digital research products is a necessary condition for reproducibility, it is not a sufficient one.
This was illustrated by an attempt to reproduce results from open materials in the journal *Cognition* [@hardwicke2018]. 
Out of 35 published articles with open code and data, results of 22 articles could be reproduced but in 11 of these cases, further assistance from the original authors was required.
For 13 articles, at least one outcome could not be reproduced—--even with the original authors' assistance.
Another study of 62 Registered Reports found that only 41 had data available, and 37 had analysis scripts available [@obels2020].
The authors could execute only 31 of the scripts without error and reproduce the results of only 21 articles (within a reasonable time).
These failed attempts to reproduce highlight the need for widely accepted reproducibility standards, because open repositories do not routinely provide sufficient information to reproduce relevant computational and statistical results.
If digital research products are available but not reproducible, their added value is limited.

Therefore, this tutorial demonstrates how to make R-based research projects reproducible, while striking a balance between rigor and ease-of-use.
A rigorous standard increases the likelihood that a project will remain reproducible as long as possible.
An easy-to-use standard, on the other hand, is more likely to be adopted.
Our approach is to promote broad adoption of easy-to-use practices, i.e., a "low threshold", and ensure that these practices are compatible with more complex rigorous solutions, i.e., a "high ceiling".
As researchers become more proficient with the tools involved, they can thus improve the reproducibility of their work.

We have structured the tutorial with a *learning by doing* approach in mind, such that readers can follow along on their own computers.
We explicitly encourage readers to try out all R commands for themselves.
Unless stated otherwise, all code blocks are meant to be run in the statistical programming language R (tested with version `r with(R.version, paste0(major, ".", minor))`).

# Threats to reproducibility and appropriate remedies

From our own experience with various research projects, we identified the following common threats to reproducibility:

1.	*Multiple inconsistent versions of code, data, or both*; for example, the data set could have changed over time because outliers were removed at a later stage or an item was later recoded; or, the analysis code could have been modified during writing of a paper because a bug was removed at some point in time.
It may then be unclear which version of code and data was used to produce some reported set of results.
2.	*Copy-and-paste errors*; for example, results are often manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result.
3.	*Undocumented or ambiguous order of computation*; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually without documentation (e.g., executing a command manually instead of in a script; copy-pasting results from one program to another).
4.	*Ambiguous software dependencies*; for example, a given analysis may depend on a specific version of a specific software package, or rely on software that might not be available on a different computer or no longer exist at all; or, a different version of the same software may produce different results.

We have developed a workflow that achieves long-term and cross-platform computational reproducibility of scientific data analyses.
It leverages established tools and practices from software engineering and rests on four pillars that address the aforementioned causes of non-reproducibility [@Peikert2019]:

1. Version control
2. Dynamic document generation
3. Dependency tracking
4. Software management

The remainder of this section briefly explains why each of these four building blocks is needed and details their role in ensuring reproducibility.
A more extensive treatment of these tools is given in @Peikert2019.

*Version control* prevents the ambiguity that arises when multiple versions of code and data are created in parallel during the lifetime (and in particular, in the development phase) of a research project.
Version control allows a clear link between which results were generated by which version of code and data.
This is important for reproducibility because only if it is obvious which version of data and code produced which results they can be said to be reproducible.
We recommend using *Git* for version control, because of its widespread adoption in the R community.

`Git` tracks changes to all project-related files (e.g., materials, data, and code) over time.
At any stage, individual files or the entire project can be compared to, or reverted to, an earlier version.
A version-controlled project makes the loss of files unlikely and also supports remote collaboration.
`Git` is built around snapshots that represent the project state at a given point in time.
Those snapshots are called *commits* and work like a "save" action.
Ideally, each commit has a message that succinctly describes these changes.
It is good practice to make commits for concrete milestones (e.g., "Commented on Introduction", "Added SES as a covariate", "Address Reviewer 2's comment 3").
This makes it easier to revert specific changes than when multiple milestones are joined in one commit, e.g., "Changes made on 19/07/2021".
Each commit refers back to its ancestor, and they form a timeline together.
The entirety of commits (i.e. the version-controlled project) is called a repository.
In `Git`, specific snapshots of a repository can be tagged, such that one can clearly label which version of the project was used to create a preregistration, preprint, or final version of the manuscript as accepted by a journal.
`Git` has additional features beyond basic version control, such as branches to facilitate collaboration.
@vuorreCuratingResearchAssets2018 provide a more extensive treatment how `Git` functions and how to use `Git` for research.
@bryanExcuseMeYou2018 provides additional information on how to track `R Markdown` documents.
Collaborating via `Git` requires the repository to be uploaded somewhere.
We recommend GitHub to host `Git` repositories because of its popularity among R users.
GitHub has many tools that ease project management and collaboration, and these tools are valuable in scientists everyday work, but they are not central to achieving reproducibility.

Second, we rely on *dynamic document generation*.
The traditional way of writing a scientific report based on a statistical data analysis uses two separate steps conducted in two different programs.
In a word processor, the researcher writes the text, and in another software, they conduct the analysis.
Typically, results are manually copied and pasted from one software to another, a process that often produces inconsistencies [@nuijtenPrevalenceStatisticalReporting2016].

Dynamic document generation integrates both steps.
Through dynamic document generation, code becomes an integral, although usually hidden, part of the manuscript, complementing the verbal description and allowing interested readers to gain a deeper understanding of the contents [@knuthCWEBSystemStructured; @claerboutElectronicDocumentsGive1992].
*R Markdown* uses `Markdown` for formatting text and R (or other programming languages) for writing the statistical analysis.
`Markdown` is a lightweight text format in plain text with a minimal set of reserved symbols for formatting instructions.
This way, `Markdown` does not need any specialized software for editing.
It is highly fool-proof [unlike, for example, `LaTeX` @lamportLATEXDocumentPreparation1994], works well with version control systems, and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
We suggest using `Markdown` for all sorts of documents that are created in the academic context, starting from simple sketches of your ideas, to your scientific manuscripts [@R-rticles] and presentations [@revealjs], or even your résumé [@vitae].
`R Markdown` extends regular `Markdown` by allowing users to include R code chunks [in fact, arbitrary computer code @riedererChapter15Other] into a `Markdown` document.
Upon rendering the document, the code blocks are executed, and their output is dynamically inserted into the document.
This allows the creation of (conditionally) formatted text, statistical results, and figures that are guaranteed to be up-to-date because they are created every time anew as the document is rendered to its output format (e.g., presentation slides or a journal article).
@xieMarkdownCookbook2020 provides an extensive yet practical introduction to most features of `R Markdown`.

One would hope that sharing a dynamic document would allow individuals (including one's future self) to simply download and reproduce a given analysis.
While version control and dynamic document generation are becoming more common, we have argued that two more components are required and that each component alone is unlikely to guarantee reproducibility [@Peikert2019; @vanlissa2020worcs].
In practice, dependencies between project files (for example, the information what script uses which data file and what script needs to be run first) or on external software (such as system libraries or components of the programming language, such as other R packages) are frequently unmentioned or not exhaustively and unambiguously documented.

To automatically resolve dependencies between project files, we rely on *dependency tracking*.
In essence, researchers provide a collection of *computational recipes*.
A computational recipe describes how inputs are processed to create a specific output in a way that is automatically executable.
The concept of computational recipes is central to our understanding of reproducibility because it enables a unified way to automatically reproduce a project.
Similar to a collection of cooking recipes, we can have multiple products (_targets_) with different ingredients (_requirements_) and different steps of preparation (_recipes_).
In the context of scientific data analysis, the targets are typically the final scientific report (e.g., the one to be submitted to a journal) and possibly intermediate results (such as preprocessed data files, simulation results, and analysis results).
One noteworthy property of computational recipes is that they require a textual representation and the ability of a program to make sense of this textual representation.
A workflow that includes, e.g. renaming variable names by hand in a graphical spreadsheet application is therefore incompatible with automated reproducibility.

We recommend using *Make* for dependency tracking because it is language independent.
The following hypothetical example illustrates the utility of `Make` and a suitable `Makefile`.
Consider a research project that contains a script to simulate data (`simulate.R`) and a scientific report of the simulation results written in `R Markdown` (`manuscript.Rmd`).
A `Makefile` for this project could look like this:

```{bash, eval=FALSE}
manuscript.pdf: manuscript.Rmd simulated_data.csv
  Rscript -e 'rmarkdown::render("manuscript.Rmd")'

simulated_data.csv: simulate.R
  Rscript -e 'source("simulate.R")'
```

There are two targets, the final rendered report (`manuscript.pdf`, l.1) and the simulation results (`simulation_results.csv`, l.4).
Each target is followed by a colon and a list of requirements.
If a requirement is newer than the target, the recipe will be executed to rebuild the target.
If a requirement does not exist, `Make` uses a recipe to build the requirement first before building the target.
Here, if one were to build the final `manuscript.pdf` by rendering the `R Markdown` with the command shown in l.2, `Make` would check whether the file `simulation_results.csv` exists; if not, it would issue the command in l.5 to run the simulation first before rendering the manuscript.
This ensures that the simulated data are present before the manuscript is built, and that the simulation is rerun and the manuscript is rebuilt if the simulation code was changed.
`Make` therefore offers a standardized process to reproduce projects, regardless of the complexity or configuration of the project.
Note that the Workflow for Open Reproducible Code in Science (WORCS) we presented elsewhere [@vanlissa2020worcs] does not contain this element of dependency tracking explicitly but its strict structure of only containing one definite `R Markdown` still makes dependencies between files unambiguous.

```{r, include=FALSE}
container_size <- as.numeric(fs::file_size("reprotutorial.sif"))/1024^3
```

A version controlled dynamic document with dependency tracking may still rely on external software.
Troubleshooting issues specific to a particular programming language or dependent tool typically requires considerable expertise and pose a threat to reproducibility.
One solution to avoid problems caused by missing or inconsistent external software dependencies is to provide all dependent software packages and system libraries via *software management*.
One comprehensive approach to software management is containerization.
The central idea is that by
"packaging the key elements of the computational  environment  needed  to  run  the  desired  software the software much easier to use, and the results easier to reproduce" [@silverSoftwareSimplified2017, p. 174].

*`Docker`* is a popular tool for containerization.
It manages software dependencies by constructing a virtual software environment independent of the host software environment.
These so-called "`Docker` images" function like a virtual computer  (i.e., a sort of "sand box").
A `Docker` image does contain *all* software dependencies used in an analysis---not just R-packages, but also R and Rstudio, and even the operating system.
All differences in computational results that could be caused by software are hence eliminated.
This is important because some of its functionality may impact the workings of higher-order software like R, such as calls to random number generators or linear algebra libraries.

Note that the software environment of the Docker image is completely separate from software installed on your computer.
This separation is excellent for reproducibility, but takes some getting used to.
For example, it is important to realize that software available on your local computer will *not* be accessible within the confines of the `Docker` image.
Each dependency that you want to use within the `Docker` image must be explicitly added as a dependency.
Furthermore, using `Docker` may require you to install software on an operating system that may not be familiar to you.
The images supplied by @boettigerIntroductionRockerDocker2017, for example, are based on Linux.

There are two ways to build a `Docker` image.
First users can manually install whatever software they like from within the virtual environment.
Such manually build environment can still be ported to all computers which support `Docker`.
We, however, prefer the second way of building images automatically from a textual description called `Dockerfile`.
Because the `Dockerfile` clearly describes which software is installed how, users can quickly change the software environment e.g. update to another R or package version without being "stuck" with an unmaintainable because manually build `Docker` image.

Packaging all required software in such an image requires considerable amounts of storage space.
Two major strategies help to keep the storage requirements reasonable.
One strategy is to rely on pre-made images, which are maintained by a community for particular purposes.
For example, there are pre-made images that only include what is necessary for R, based on Ubuntu containers [@boettigerIntroductionRockerDocker2017].
Users can then install whatever they need in addition to what is provided by these pre-compiled images.
`r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disk space."))`
The image for this project includes Ubuntu, R, RStudio, LaTeX as well as a variety of R packages like tidyverse [@tidyverse] and all its dependent packages, amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` R packages.
A second strategy is to save a container recipe; a so-called  `Dockerfile` containing a textual description of all commands that need to be executed to recreate the software environment, instead of the software environment itself.
`Dockerfiles` are tiny (the [`Dockerfile`](https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile) for this project has a size of only `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb).
However, they rely on the assumption that all software repositories that provide the dependent operating systems, pieces of software, and R packages will remain accessible in future and will continue to make available historic software versions.
For proper archival, we therefore recommend to store a complete image of the software environment, in addition to the `Dockerfile`.
A more comprehensive overview of the use of containarization in research projects is given by @wiebelsLeveragingContainersReproducible2021.
Note that the Workflow for Open Reproducible Code in Science (WORCS) we presented elsewhere [@vanlissa2020worcs] relies for software management on the R package `renv`[@R-renv] which is more lightweight and easier to use than `Docker` on the one hand, but not as comprehensive on the other hand because it only snapshots the R packages instead of all software used.

To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)), dynamic document generation (using `R Markdown`), version control (using `Git`), dependency tracking (using `Make`), and Software management (using `Docker`).
While `R Markdown` and `Git` are well integrated into the R environment through RStudio, `Make` and `Docker` require a level of expertise that is often beyond the training of scholars outside the field of information technology, which can represent a considerable obstacle to the acceptance and implementation of the workflow.
To remove this obstacle, we have developed the R package `repro` that supports scholars in setting up, maintaining, and reproducing research projects in R.
Importantly, a reproducible research project created with `repro` does not have the `repro` package itself as a dependency.
These projects will remain reproducible irrespective of whether `repro` remains accessible in future.
Users do not have to have `repro` installed to reproduce a project;
in fact, they do not even need to have R installed because the entire project can be rebuilt inside of a container that has R installed.
In the remainder, we will walk you through the creation of a reproducible research project with the package `repro`.

```{r schematic, eval = TRUE, echo = FALSE, fig.cap="Schematic illustration of the interplay of the four components (in dashed boxes) central to the reproducible workflow: version control (`Git`), dependency tracking (`Make`), software management (`Docker`), and dynamic document generation (`R Markdown`). `Git` tracks changes to the project over time. `Make` manages dependencies among the files. `Docker` provides a container in which the final report is built using dynamic document generation in `R Markdown`. Reproduced from @Peikert2019 licensed under [CC BY 4.0](https://creativecommons.org/licenses/by/4.0).  "}
# file gets downloaded in Makefile
knitr::include_graphics(here::here("images/nutshell.svg"), auto_pdf = TRUE)
```

# Creating reproducible research projects

One impediment to the widespread adoption of a standard for reproducible research is that learning to use the required tools can be quite time-intensive.
To lower the threshold to adoption, the R-package `repro` introduces helper functions that wraps and hides the complexity of the necessary more complicated (and powerful) tools.
The provided wrapper functions guide end-users in the use of reproducibility tools, provide feedback about what the computer does, and suggest what the user is supposed to do next.
The `repro` package extends the [`usethis`](https://usethis.r-lib.org) [@usethis] package.
While the `usethis` package aims to simplifies the development of R packages, `repro` contains reproducibility specific utilities.
We hope this makes reproducibility tools more accessible by enabling beginner-level users to detect their system's state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
These wrappers are merely an assistance system, and as users get more and more comfortable, they can use the underlying tools directly to solve more complex problems.

This tutorial assumes that the user is working predominantly in R, with the help of RStudio.
It describes basic steps that we expect to be representative of a small-scale psychological research project. 
Of course, your specific situation might involve additional, more specialized steps. 
After completing the tutorial, you should be able to customize your workflow accordingly.
The first step is to install the required software.
We assume that you have installed R [@R-base, version `r with(R.version, paste0(major, ".", minor))`] and RStudio [@rstudio, version 1.4] already but the tutorial will guide you in detail through the installation of other necessary software with the help of the R package [`repro`](https://github.com/aaronpeikert)[@R-repro].
In case you either have not installed R and RStudio or are unsure if they are up to date you might want to consult our installation advice in the [Online Supplementary Material](https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md) that covers how to install all software necessary for this tutorial in three steps.
The [installation advice]((https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md)) may also help Windows users who have problems installing `Docker`.
Start RStudio and install the package [`repro`](https://github.com/aaronpeikert)[@R-repro]. 
It will assist you while you follow the tutorial.

```{r, eval=FALSE}
# repro is not on CRAN yet
options(
  repos = c(aaronpeikert = 'https://aaronpeikert.r-universe.dev',
            CRAN = 'https://cloud.r-project.org')
)
install.packages('repro')
```

An example of the assistance that `repro` provides are the check functions.
These verify that you indeed have installed and setup the required pieces of software for this workflow (here, we use the double-colon operator to access functions from the `repro` namespace; we chose to do so to make clear which functions are native `repro` functions) :

```{r, include=FALSE}
opts <- options()
options(repro.docker = TRUE,
        repro.docker.running = FALSE)
```

```{r}
# `package::function()` → use function from package without `library(package)`
repro::check_git()
repro::check_make()
repro::check_docker()
```

```{r, include=FALSE}
options(opts)
```

These functions check whether specific dependencies are available on the users' system, and if not, explain what further action is needed to obtain it.
Sometimes they ask the user to do something, for example, the following happens if you were a Windows user who does not have `Git` installed:

```{r, include=FALSE}
opts <- options()
options(repro.os = "windows",
        repro.git = FALSE)
```

```{r}
repro::check_git()
```

```{r, include=FALSE}
options(opts)
```

The messages from `repro` try to take the user by the hand to help them solve problems.
These messages are adjusted to your specific operating system and installed dependencies.
Before you continue, we ask you to run the above commands to check `Git`, `Make`, and `Docker`---both to become familiar with the functionality of the `check_*()` functions, and to prepare your system for creating reproducible projects.
If necessary, follow any instructions presented until all checks pass.

## Implementing version control

Now that your system is set up for this tutorial, we will introduce you to version control with `Git`.
`Git` tracks files within one project folder.
We start by creating such a project folder with RStudio by clicking the menu item:

> File → New Project... → New Directory → Example Repro Template

This creates a project containing an example analysis, but you may use any other template or existing R Project to turn into a reproducible research project using the functionality of `repro`.

`Git` tracks any files that you manually add to the repository.
To make sure you do not accidentally add files that you do not wish to share (e.g., privacy sensitive data),
you can list specific files that you do not want to track in the `.gitignore` file.
You can add something to the `.gitignore` file directly or with this command:

```{r, eval=FALSE}
usethis::use_git_ignore("private.md")
```
Now the file `private.md` will not be committed to `Git` and will not be made public.

Especially for new users, it is recommended to follow this procedure and exclude sensitive files before proceeding.
When you are ready, you can begin tracking your files using `Git` by running:

```{r, eval=FALSE}
usethis::use_git()
```

For `Git` to recognize changes to a given file, you have to stage and then commit these changes (this is the basic save action for a project snapshot).
One way to do this is through the visual user interface in the RStudio `Git` pane.
Click on the empty box next to the file you want to stage.
A check mark appears to indicate that the file is staged.
After you have staged all files you want, click on the commit button, explain in the commit message why you made those changes, and then click on commit.
This stores a snapshot of the current state of the project.

The files you created and the changes you made still have not left your computer.
All snapshots are stored in a local repository in your project folder.
To back up and/or share the files online, you can push your local repository to a remote repository.
While you can choose any `Git` service (like GitLab or BitBucket), we will use GitHub in this tutorial.
Before you upload your project to `Git`, you need to decide if you would like the project publically accessable (viewable by anyone, editable by selected collaborators) or if you want to keep it private (only viewable and editable by selected collaborators).
To upload the project publically to GitHub use:

```{r, eval=FALSE}
usethis::use_github()
```

To upload it privately:

```{r, eval=FALSE}
usethis::use_github(private = TRUE)
```

Depending on your computer configuration, this may ask to set up a secure connection to GitHub.
In this case first follow the suggestions shown on the R console.

## Using dynamic document generation

Now that you have created a version controlled project, we will proceed with dynamic document generation.
A dynamic document has three elements:

1. Text (prose; e.g., a scientific paper or presentation)
2. Executable code (e.g., analyses)
3. Metadata (e.g. title, authors, document format)

`R Markdown` is a type of dynamic document well-suited to the RStudio user interface.
The text of an `R Markdown` is formatted by `Markdown` (see @xieMarkdownDefinitiveGuide2019 for technical details and @xieMarkdownCookbook2020 for practical guidance).
The code mostly consists of R-code (although other programming languages are supported, like Python, C++, Fortran, etc).
The following example serves to illustrate the `Markdown` syntax. 
It shows how to create a heading, a word in bold font, a citation and a list of several items in `Markdown`:

```markdown
<!--this is a Markdown file -->
# Heading (level 1)

Normal text.
Important **word** in bold.
A citation: @einstein1935 did important research on this topic.

## Subheading (level 2)

To do list:

* do research
* do more research
* spend time with family and friends
```

One advantage of this type of markup for formatting is that it can be rendered to many different output formats---both in terms of file types, like `.docx`, `.html`, `.pdf`, and in terms of style, e.g. specific journal requirements.
Another advantage is that, because `R Markdown` files are in plain text, they are more suitable for version control using `Git`.
Some users might find it easier to activate the "Visual Editor" of RStudio (Ctrl + Shift + F4 or click on the icon that resembles drawing materials or a compass in the upper right corner of the `R Markdown` document) which features more graphical elements like a traditional word processor but still creates an `R Markdown` underneath with all of its flexibility.
The visual editor has some additional benefits, such as promoting best practices (for example, each sentence should be written on a newline, which eases tracking changes across versions)
and easing the generation of citations and references to Tables and Figures.

Now that you are familiar with the basics of `Markdown` formatting, we turn our attention to including code into the text.
Code is separated by three backticks (and the programming language in curly braces) like here:

````markdown
This is normal text, written in Markdown.

`r ''````{r}
# this is R code
1 + 1
```
````

The hotkey [Control]+[Alt]+[i] inserts a block of code in the file.
The results of code enclosed in such backticks will be dynamically inserted into the document (depending on specific settings).
This means that, whenever you render the `R Markdown` to its intended output format, the code will be executed and the results updated.
The resulting output document will be static, e.g. a pdf document, and can be shared wherever you like, e.g on a preprint server.

Once the `R Markdown` file has been rendered to a static document, the resulting file is decoupled from the `R Markdown` and the code that created it.
This introduces a risk that multiple versions of the static document are disseminated, each with slightly different results.
To avoid ambiguity, we therefore recommend referencing the identifier of the `Git` commit at the time of rendering in the static document.
Simply put, a static document should link to the version of the code that was used to create it.
The `repro` package comes with the function `repro::current_hash()` for this purpose.
This document was created from the commit with the hash [``r repro::current_hash()`` (view on GitHub)](`r stringr::str_c("https://github.com/aaronpeikert/repro-tutorial/tree/", repro::current_hash(40))`).

Now that you know how to write text and R code in an `R Markdown`, you need to know about metadata (also called: YAML front matter).
These metadata contain information about the document, like the title and the output format.
Metadata are placed at the beginning of the document and are separated by three dashes from the document body.
The following example is a full markdown document where the metadata (the "YAML front matter") is in lines 1--6.

````markdown
---
title: "A tutorial on how to do the same thing more than once"
author: Aaron Peikert, Caspar J. Van Lissa and Andreas M. Brandmaier
abstract: A Hitchhiker's Guide to Reproducible Research in R
output: html_document
---

# Introduction

Important for reproducibility:

1. *Version Control*
2. *Dynamic Document Creation*
3. *Dependency Tracking*
4. *Software Management*

`r ''````{r}
# this is R code
t.test(extra ~ group, data = sleep)
```
````

## Manage software and file dependencies

Some metadata fields are self-explanatory (like the author field), and exist across all output formats (like the title field).
However, `R Markdown` has been extended to hundreds of document formats which are often highly configurable by their metadata.
Some R packages add additional metadata fields.
The `repro` package adds fields to the metadata to help make the research project reproducible.
These fields are used to list all dependencies of the research project.
This includes R scripts, data files, and external packages.
The format is as follows (see everything below the line *repro:*):

```markdown
---
title: "A tutorial on how to do the same thing more than once"
author: Aaron Peikert, Caspar J. Van Lissa and Andreas M. Brandmaier
output: html_document
repro:
  scripts:
    - R/load.R
  data:
    - data/mtcars.csv
  packages:
    - tidyverse
    - usethis
    - gert
---
```

This information makes explicit what dependencies (in form of files and R packages) a project relies on .
`repro` uses this information to construct a `Makefile` for the dependencies on other files and a `Dockerfile` that includes all required packages.
These two files together form the basis for consistency within a research project and consistency across different systems.
The function `repro::automate()` converts the metadata from all `R Markdown` files in the project (all files with the ending `.Rmd`) to a `Makefile` and a `Dockerfile`.
Together, these files allow users (including your future self) to reproduce every step in the analysis automatically.
It is important to re-run `repro::automate()` whenever you change the `repro`-metadata, change the output format, or add a new `R Markdown` file to the project, to keep the `Makefile` and `Dockerfile` up to date.
There is no harm in running it too often.
Other than the `Makefile` and the `Dockerfile`, which are created in the document root path, repro generates a few more files in the `.repro` directory (which we will explain in detail later), all of which you should add and commit to `Git`.

## Reproducing a project

When you or someone else who received your project want to reproduce it, they first have to install the required software, which is `Make`, and `Docker`.
Remember, you can use the `check_*`-functions to test if these are installed:

```{r}
repro::check_make()
repro::check_docker()
```

When they are setup, they can ask `repro` to explain how they should use `Make` and `Docker` to reproduce the project (or you could explain it to them):

```{r}
repro::reproduce()
```

If you feel uncomfortable using the terminal directly, you can sent the command to the terminal from within R:

```{r, eval=FALSE}
system(repro::reproduce())
```

The only software required for reproducing a project is `Docker`, assuming users know how to build a `Docker` image and run `Make` within the container.
However, if they have installed `Make` additionally to `Docker`, they do not need to know how to use `Docker` and can simply rely on the two `Make` commands `make docker` and `make -B DOCKER=TRUE`.
The commands can be pasted into a terminal and will reproduce the project in many cases.

## Summary

1. Install the `repro` package:

```{r, eval=FALSE}
options(
  repos = c(aaronpeikert = 'https://aaronpeikert.r-universe.dev',
            CRAN = 'https://cloud.r-project.org')
)
install.packages('repro')
```

2. Check the required software:

```{r}
repro::check_git()
repro::check_make()
repro::check_docker()
```

3. Create an R project or use an existing one. Do not forget to add `repro` metadata (i.e. packages, scripts, data).

```yaml
repro:
  scripts:
    - R/load.R
  data:
    - data/mtcars.csv
  packages:
    - tidyverse
```

The example `repro` project already has this metadata:

```{r,eval=FALSE}
repro::use_repro_template("/some/folder")
```

4. Let `repro` generate `Docker`- and `Makefile`:

```{r, eval=FALSE}
repro::automate()
```

5. Enjoy automated reproducibility:

```{r}
repro::reproduce()
```

# Advanced Features

This section is only for advanced users who want to overcome some limitations of `repro`.
If you read this paper the first time you probably want to skip this section and continue reading at the section "Preregistration as Code".
As explained above, `repro` is merely a simplified interface that enables a high level of reproducibility for simple projects.
The simplification of `Docker` and `Make` are possible because of two restrictions.
Users who ask themselves either "How can I install software dependencies outside of R in the `Docker` image?" or "How can I express complex dependencies between files, e.g. hundreds of data files are preprocessed and combined?") need to be aware of these restriction and require a deeper understanding of the inner workings of `repro`.
Other users may safely skip this section, or return to it if they encounter such challenges.

The metadata the `repro::automate()` function relies on can only express R packages as dependencies for the `Dockerfile` and only trivial dependencies (in the form of "file must exist") for the `Makefile`. 
The first limitation means that users must rely on software that is either already provided by the base `Dockerimage` "rocker/verse" or the R packages they list in the metadata.
Other software that users might need like other programming languages, not already installed LaTeX packages etc. have to be added manually.
We plan to add support for commonly used ways to install software beyond R packages via the metadata and `repro::automate()` e.g. for system libraries (via `apt` the Ubuntu package manager), LaTeX packages (via `tlmgr` the Tex Live package manager), python packages (via `pip` the python package manager).
The second limitation is about dependencies.
`Make` can represent complex dependencies like in this example: A depends on B, which in turn depends on C and D.
If B is missing in this example, `Make` would know how to recreate it from C and D.
These dependencies and especially how these dependencies should be resolved are difficult to represent in the metadata.
Users therefore have to either "flatten" the dependency structure, by simply stating that A depends on B, C and D, therefore leaving out important information or express the dependencies directly within the `Makefile`.

The following section explains how to overcome these limitations while still relying on the automations afforded by `repro`.
Lifting these restrictions requires the user to interact more directly with `Make` or `Docker`.
To satisfy more complicated requirements users need to understand how repro utilizes `Make` and `Docker` internally.

Let us have a closer look at the command for reproducing a repro project: `make docker && make -B DOCKER=TRUE`; it consists of two processing steps.
First it recreates the virtual software environment (`Docker`) and then it executes computational recipes in the virtual software environment (`Make`).
The first step is done by the command `make docker`.
The command `make docker` will trigger `Make` to build the target called `docker`.
The recipe for this target builds an image from the `Dockerfile` of the repository.
The `&&` concatenates both commands and only runs the second command if the first was successful.
Therefore, the computational steps are only executed when the software environment is set up.
The second step executes the actual reproduction and is again a call to `Make` in the form of `make -B DOCKER=TRUE` with three noteworthy parts.
First, a call to `make` without any explicit target will build the `Make` target `all`.
Second, the flag `-B` means that `Make` will consider all dependencies as outdated and hence will rebuild everything.
Third, repro constructs `Make` targets so that if you supply `DOCKER=TRUE` they are executed within the `Docker` image of the project.

This interplay between `Docker` and `Make` is relatively complicated because it resembles a chicken or egg problem.
We have computational steps (`Make`) which depend on the software environment (`Docker`) for which we again have computational steps that create it.
Users only require a deeper understanding of this interdependence when they either want to have more complex computational recipes than rendering an `R Markdown` or require other software than R packages.

Users can have full control over the software installed within the image of the project.
`repro` creates three `Dockerfiles` inside the `.repro` directory.
Two `Dockerfiles` are automatically generated.
The first is `.repro/Dockerfile_base`.
It contains information about the base image on which all the remaining software is installed.
By default we rely on the "verse" images provided by the Rocker project [@boettigerIntroductionRockerDocker2017].
These contain (among other software) the packages `tidyverse`, `rmarkdown` and a complete LaTeX installation, which makes these images ideal for the creation of scientific manuscripts.
Users can choose which R version they want to have inside the container by changing the version number in line 1 to the desired R version number.
By default the R version corresponds to the locally installed version on which `repro::automate()` was called the first time.
The build date is used to install packages in the version which was available on CRAN on this specific date and can also be changed, by default this date is set to the date on which `repro::automate()` was called the first time. 
This way, the call to the automate function virtually freezes the R environment to the state it was called the first time inside the container. 
Below, you see the `Docker` base file we used to create this manuscript:

```{r, results='asis', echo=FALSE}
cat(
  "```bash",
  readLines(here::here(".repro", "Dockerfile_base"))[1:3],
  "```",
  sep = "\n"
)
```

The second automatically generated `Dockerfile` is `.repro/Dockerfile_packages`.
Whenever `repro::automate()` is called, repro gathers all R packages from all `.Rmd` files and figures out whether they should be installed from CRAN or GitHub fixed to the date specified in `Dockerfile_base`.

Finally, there is one manually edited `Dockerfile`: `.repro/Dockerfile_manual`.
It is blank by default, and can be used to add further dependencies outside of R, like system libraries or external software.

Using `Docker` may require you to install software on an operating system that may not be familiar to you.
The images supplied by [@boettigerIntroductionRockerDocker2017], for example, are based on the Ubuntu operating system.
The most convenient way to install software on Ubuntu is through its package manager `apt`.

If the following snippet is added to `.repro/Dockerfile_manual` the `Docker` image will have, for example, Python installed.
Other software is installed identically only the software name is exchanged.

```bash
RUN apt-get update && apt-get install -y python3
```

`Docker` eventually requires a single `Dockerfile` to run, so `repro::automate()` simply concatenates the three `Dockerfiles` and saves the result into the main `Dockerfile` at the top level of the R project.

With this approach, users of `repro` can not only build complex software environments but they can also implement complex file dependencies.
The standard `repro` metadata only makes sure that all dependencies are available but does not allow you to specify custom recipes for them in the metadata.
If you can formulate the creation of dependencies in terms of computational steps, e.g. the file `data/clean.csv` is created from `data/raw.csv` by script `R/preprocess.R`, you should include these in the `Makefile`.
The `Makefile` that `repro` creates is only a template and you are free to change it.
However, make sure to never remove the following two lines:

```bash
include .repro/Makefile_Rmds
include .repro/Makefile_Docker
```

The file `.repro/Makefile_Rmds` contains the automatically generated targets from `repro::automate()` for the `R Markdown` files.
This file should not be altered manually.
If you are not satisfied with the automatically generated target simply provide an alternative target in the main `Makefile`.
Targets in the main `Makefile` take precedent.

The file `.repro/Makefile_Docker` does again contain a rather complicated template which you could but usually should not modify.
This `Makefile` coordinates the interplay between `Make` and `Docker` and contains targets for building (with `make docker`) and saving (with `make save-docker`) the `Docker` image.
Additionally, it provides facilities to execute commands within the container.
If you write a computational recipe for a target it will be by default evaluated using the locally installed software.
To instead evaluate commands inside the `Docker` image, you should wrap them in `$(RUN1) command $(RUN2)`, like in this example, which is identical to the first `Make` example we gave above except for the addition of `$(RUN1)` and `$(RUN2)` in l. 2:
```bash
simulated_data.csv: R/simulate.R
  $(RUN1) Rscript -e 'source("R/simulate.R")' $(RUN1)
```

If users execute this in the terminal:

```bash
make data/simulation_results.csv
```
It behaves exactly as the in the first `Make` example, the script `R/simulate.R` is run using the locally installed R.
Because this translates simply to:

```{r, echo=FALSE, results='asis'}
cat("```bash",
  stringi::stri_wrap(system2("make", " --dry-run data/simulation_results.csv", stdout = TRUE)),
  "```", sep = "\n")
```

But if users use

```bash
make DOCKER=TRUE data/simulation_results.csv
```

it is evaluated within the `Docker` container using the software within it and not the locally installed R version:

```{r, echo=FALSE, results='asis'}
cat("```bash",
  stringi::stri_wrap(system2("make", " --dry-run DOCKER=TRUE simulated_data.csv", stdout = TRUE)),
  "```", sep = "\n")
```

To summarise, `repro` provides researchers with an easy to use solution for dependency tracking (in form of `Make`) and software management (using `Docker`) without the necessity to learn both tools.
Users with more advanced requirements can manually change all aspects of both programs while still profiting from `repro`'s automations.

# Preregistration as Code

Defining the research questions and planning data analysis before observing the research outcomes is referred to as *preregistration* [@NosekRevolution2018].
Preregistration increases the credibility of empirical results by separating a-priori planned and theory-driven analyses from those conducted unplanned and post-hoc, that is, after having seen the data.
Preregistration seeks to reduce "researcher degrees of freedom" [@NosekRevolution2018] in the execution of a planned study by limiting the number of decisions to be made after data has been collected, which could bias the reported results.
Preregistrations, however, have several shortcomings.
One shortcoming is that written study plans are often multi-interpretable and often do not achieve the level of unambiguity that are required to effectively zero the researcher degrees of freedom.
Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis, the results vary considerably [@silberzahnManyAnalystsOne2018].
The current best practice to ensure that the preregistration are comprehensive and specific enough is to use preregistration templates [like @bowmanOSFPreregTemplate2020].
These templates seek to improve the specificity of study plans by imposing structure [@bakkerEnsuringQualitySpecificity2020].
However, templates cannot remove researchers degrees of freedoms completely, because it is impossible to verbally describe every detail of an analysis for all but the most straightforward analysis.
This ambiguity causes a second problem, that is comparing the preregistration and the resulting publication to decide if and how researcher deviated from the preregistration.
This task is difficult because it impossible to decide without additional information if the analysis actually was carried out differently or just described differently.
Even when the researcher was faithful in each aspect to the preregistration the reader may reach opposite conclusions, because the reader has to compare to different texts which may be worded differently or describe the same thing in varying level of detail.
A third limitation is that, just like primary research, preregistrations are susceptible to non-reproducibility.
To illustrate, a review of 210 preregistrations found that, even though 174 (67%) included a formal power analysis, only 34 (20%) of these could be reproduced [@bakkerRecommendationsPreregistrationsInternal2020].
Even when researchers have gone to great length in preregistering an analysis script,
they sometimes inexplicably fail to reproduce their own results.
For example, @steegenMeasuringCrowdAgain2014 realized after publication that part of their preregistered code results in different test statistic than they reported initially (see Footnote 7).
A final limitation is that written plans may turn out to be unfeasible once data are obtained and analyzed.
For example, a verbal description of a statistical model may be unidentified, e.g. if it includes reciprocal paths between variables, or more parameters than observed data.
Conversely, a model may be mis-specified in a major way; for example by omitting direct effects when the research question is about mediation, thus leading to a model with unacceptable fit.
Many researchers would only realize that such a model cannot be estimated once the data are obtained, thus necessitating a deviation from the preregistered plans.

The workflow described in this paper facilitates a rigorous solution to this problem:
Instead of describing the analysis in prose, researchers include the code required to conduct the analysis in the preregistration.
We term this approach of writing and publishing code at the preregistration stage *preregistration as code (PAC)*.
PAC has the potential to eliminate researchers degrees of freedom to a much greater extent than, e.g., preregistration templates.
Moreover, it removes the necessity to write two separate documents:
A preregistration *and* a manuscript.
For a PAC, researchers can write a reproducible, dynamically generated draft of their intended manuscript at the preregistration stage.
This already includes most typical sections, such as Introduction, Methods, and Results.
These results are initially based on simulated data with the same structure as the data they expect to obtain from their experiments.
For guidance on how to simulate data, see @morrisUsingSimulationStudies2019, @paxtonMonteCarloExperiments2001 and @skrondalDesignAnalysisMonte2000.
Once the preregistration is submitted and real data have been collected or made available,
the document can be reproduced with a single command, thus updating the Results to their final version.

The idea of submitting code as part of a preregistration is not new.
A prominent preregistration platform, [The Open Science Framework](https://osf.io/), suggests submitting scripts alongside the preregistration of methods.
In an informal literature search (we skimmed the first 300 results with the keywords `("pre registration"|"pre-registration"|preregistration)&(code|script|matlab|python|"R")`) we have found close to a dozen published articles which did include some form of script as part of their preregistration.
But although the notion of preregistering code has been around for a while [cf. @steegenMeasuringCrowdAgain2014],
it has not gained much traction---perhaps because, until now, this has constituted an extra non-standard step in the research process.
This tutorial integrates the preregistration of code into the reproducible research workflow
by encouraging researchers to preregister the whole manuscript as a dynamic document.

## Advantages of PAC over traditional preregistration

We believe that pairing PAC with the here presented workflow offers five advantages over classical preregistration.
First, the PAC is merely an intermediate stage of the final manuscript, thus sparing authors from writing, and editors and reviewers from evaluating two separate documents.
Relatedly, writing a preregistration in form of a research article has the advantage that researchers already have great expertise in writing a scientific manuscript.
By contrast, filling out a preregistration template is a novel type of scientific writing for many.
Second, the PAC is a tool for study planning.
When all steps to be executed are documented clearly, the execution of the study will proceed more efficiently than when every step is planned in medias res.
Third, a PAC removes any ambiguity regarding the translation of verbal analysis plans into code.
A PAC is more comprehensive by design because its completeness can be empirically checked with simulated data.
Evaluating the intended analysis code on simulated data will help identify missing steps or ambiguous decisions.
PAC therefore minimizes researchers degrees of freedom more effectively than a standard preregistration [@bakkerEnsuringQualitySpecificity2020].
Fourth, despite its rigor, PAC does accommodate data-dependent decisions if these can be formulated as code.
Researchers can, for example, formulate conditions (e.g., in the form of if-else-blocks) under which they prefer one analysis type over the other.
For example, if distributional assumptions are not met, the code may branch out to employ robust methods; or, an analysis may perform automated variable selection mechanisms before running the final model.
Another example of data dependent decisions are more explorative analyses, i.e. explorative factor analysis or machine learning.
With PAC researchers can precisely calibrate how much they want to explore.
When communicating a finding as explorative, it may still make a difference if tens or hundreds of variables were considered.
Fifth, deviations from the preregistration are clearly documented because they are reflected in changes to the code, which are managed and tracked with version control.

## Deviating from the preregistration and exploration

The convenient comparison of preregistration and publication with PAC is particularly noteworthy.
Because there is a single "source of truth", the dynamic document, including code, under version control, changes can be inspected by readers and authors alike.
To make the comparison more accessible, we highly recommend that the authors review all changes to summarize and justify them for their reader, e.g. in the discussion [^summarize-changes].
The reader is also in a position to inspects the changes and can therefore make an informed judgment about accepting or rejecting the justifications.

Deviations from the preregistration are sometimes demonized as if encountering unexpected aspects would invalidate a whole carefully crafted study.
We, however, oppose this view.
The problem is not deviating from the preregistration; it is not making these changes transparent.
In fact, we expect that most PACs will require changes after the researcher collects data.
Often deviations provide an opportunity to learn from the unexpected.
Transparency is essential in this regard because an empirical test of a theory can not test the theory alone but tests it in conjunction with auxiliary assumption [@meehlTheoreticalRisksTabular1978].

Take the following example: A preregistration intended to include "working memory" and "fluid intelligence" as predictors for reaction time in an experiment.
However, after fitting the model on the actual data, both constructs exhibit high collinearity.
The authors decide that for their purpose, these variables are interchangeable and remove one.
For them, this change simply does not affect their theory's core but was an auxiliary assumption, and they might be right.
Another researcher, however, who, e.g. hopes to incorporate exactly this result into a theory about intelligence, might disagree about the importance of this assumption and would accordingly view the finding as more explorative because of the change to the preregistrations.
In other words, "one man's trash is another man's treasure".

[^summarize-changes]: In section [Preregistration as Code --- a Tutorial] we have conducted an actual PAC and summarize the changes we made to the preregistered code in the discussion.

Transparency of deviations is also preferable to avoiding them by vagueness in the first place.
A vague preregistration must have reduced diagnosticity for demarcating exploration and conformation.
A reader may falsely conclude that a result is explorative because the publication is more specific than the preregistration, therefore committing an alpha error (believing there is a deviation when, in fact, there is none).
At the same time, a researcher may overlook a deviation---the goal of intentional vagueness---therefore committing a beta error.
At its most extreme, the preregistration "We will do something with some data." has lost all ability to differentiate between exploration and conformation, which was what we set out to do with preregistrations.

Another common misunderstanding is that preregistration, and therefore PAC, preclude exploratory analyses.
We must differentiate here between two kinds of exploration, and neither of them is inhibited.
The first and more traditional meaning refers to ad hoc statistical procedures and post hoc explanations of the results.
Such traditional exploratory results should merely be explicitly declared in the manuscript to allow appropriate judgment of the results.
The second kind of exploration is through procedurally well-defined exploration with exploratory statistical models, which are standard in machine learning.
These highly-parameterized models, which are difficult to preregister traditionally, can be preregistered with PAC like any other model.
We recommend PAC for those exploratory statistical models without any irony.
The merit of preregistration in such cases is to communicate precisely how much exploration was done, a piece of information that is crucial for assessing the uncertainty of such now "preregisterable" but still exploratory results.

## What you need to know before writing a PAC

When planning a study and hence when writing a PAC, researchers ought to strive to meet three goals:
First, the plan should be comprehensive, that is, researchers need to state the research question and describe the study design and analysis in detail.
Second, the plan should be effective, meaning that the researchers can reasonably expect it to answer the research question at hand.
Third, the plan should be efficient; the plan should only require the necessary amount of resources.

### Comprehensiveness

Standards for scientific writing are designed to result in understandable and comprehensive research articles and can serve the same purpose for writing a preregistration.
Standards for scientific writing [e.g. @apa7] require that research articles are described in such detail that any researcher that is familiar with the field is able to repeat a given study.
They provide, hence, an excellent vehicle for planning a study.
One particular advantage is the foundation in existing literature where the introduction and theoretical background provide the basis for the studies' design.
Nevertheless, it can be fruitful to consult preregistration templates because of their strong focus on formulating a comprehensive methods section.
If the PAC is submitted as a registered report [an article format in which the preregistration is reviewed and the article accepted "in principle", @nosekRegisteredReports2014; @chambersWhatNextRegistered2019] a journal article might be more familiar to the reviewers that ensure a comprehensive study plan.

Another factor that ensures a comprehensive plan is writing the analysis based on simulated data.
Simulating data offers various benefits in the planning phase of a study.
It forces the researcher to think about and address data analytic problems on a conceptual level before they make the effort to acquire data [@axelrodAdvancingArtSimulation1997].
It also affords researchers with a simple technical check whether the analysis code runs without failure [@braiekTestingMachineLearning2020] (for example, does the code work with missing values, extreme outliers, etc.).
Writing incomplete code is technically impossible, or at least researchers would notice if something is omitted (e.g. if they forgot to aggregate the data).

### Effectiveness

When the comprehensiveness criterion is met, researchers can check if the analysis is effective at recovering the true parameters of the simulation.
If an analysis fails to recover the correct values in idealized (simulated) conditions,
it would be unreasonable to expect this analysis to work under realistic conditions.
Because the simulation describes the researchers expectations about real world conditions,
it is the best way to assess the effectiveness of the analysis without actually gathering the data.
A simulation of data requires the researchers to translate their expectations about the underlying process into concrete computational terms, e.g. assumptions about which distribution does the dependent variable follow, or how do the independent variables influence each other and the dependent variable.
If a researcher experiences difficulty in specifying such expectations, that may be an indication that further literature research, or exploratory pilot studies, are required.

### Efficiency

A single simulated dataset already helps to ensure that the plan is viable under certain circumstances.
However, repeated simulation with varying parameters, a so-called Monte Carlo simulation, lets researchers play through several scenarios.
Monte Carlo simulation lets researchers navigate several trade-offs regarding their resources.
In the simplest case, they can assess the power (the probability to detect an effect of a given size) for varying numbers of observations in the simulation to determine the required sample size for their study [@brandmaier2018precision].
@bakkerRecommendationsPreregistrationsInternal2020 found that a formal power analysis does not increase the planned sample size on avaerage but could help to invest resources more efficiently.
Researchers may also evaluate how the data analysis copes with expected data complications, e.g. non-normal data [@harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].

### The writing process

To write a PAC, researchers should start by drafting a manuscript in the form of an `R Markdown`.
When they have have outlined the studies design, they should begin writing code.
We suggest writing three functions for each planned hypothesis, one for the planned analysis, one for simulating the data and one for reporting the results.
We capture these distinct tasks in the form of explicit R-functions but researcher may use any form (e.g. scripts) as long as it serves to simulate, analyze, and report.
The simulate and report function can of course be reused, if appropriate, across multiple hypotheses.
At the heart of a PAC lies the planned analysis, so researchers should first draft a function called `planned_analysis()` from which they later derive the method section.
This function should receive the data and compute all relevant results from it.

```{r}
planned_analyis <- function(data){
  # 1. preprocess e.g. with `rowMeans(data)`
  # 2. conduct analysis e.g. with `t.test()`
  # 3. `return(results)`
}
```

It is difficult to write code when it is not clear how exactly the input should look like, however, at the time of preregistration, researchers should remain insulated from real data.
We therefore simulate a dataset first that resembles the expected structure of the empirical data that we will use for the final analysis.
Using the simulated dataset as placeholder, we can now write a scientific report including all analysis code blocks and the presentation of results, which can serve as PAC.

```{r, eval=FALSE}
simulate_data <- function(n, effect_size){
  # 1. warn users that the results are "fake"
  # 2. draw `n` samples with `effect_size`
  # 3. format and return in expected data format
}
```

As soon as they have written `planned_anaylsis()` and `simulate_data()` they can iteratively improve both functions, e.g. until `planned_analysis()` runs without error and recovers the correct parameters from `simulate_data()`.
The goal is that the output of `simulate_data()` works as input to the function `planned_analysis()`.

#### Alternatives to simulated data

If simulating data proves to be too challenging, researcher can delay their preregistration till they have gathered data and repurpose the actual data.
In order to do so researcher must walk a fine but utmost important line.
Of course researcher must not analyze the data before they write the preregistration.
Nevertheless, they may obfuscate the real data by shuffling all variables or otherwise create synthetic data that allows them to test the function `planned_analysis()` without actually revealing the research outcomes.
Preregistration after data collection is common when reanalyzing data from another research group [@westonRecommendationsIncreasingTransparency2019] but not so much within the same research project.
We argue that it is still an eligible preregistration.
Since the data is obfuscated, almost no results are revealed and the researchers can not employ their researcher degrees of freedom effectively to influence the results.
Shuffling leaves only univariate information, so this procedure is only applicable for bi-or multivariate hypotheses.
Researcher still can access the information about means or proportions (e.g. how many participants belonging to group "A" are in the dataset) but are unbiased for all relations between variables (e.g. members of group "A" have a greater mean in variable "Z").

When the researchers are satisfied with the function `planned_analysis()`, they can think about how they would like to report the analysis results via tables, plots, and text.
The implementation of this reporting should be in the function `report_analysis()`.

```{r, eval=FALSE}
report_analysis <- function(results){
  # 1. create markdown tables from results
  # 2. conditionally interpret results e.g. if(p < .025)"Result is significant."
  # (optional) visualize results
  # 3. return results section formatted in markdown
}
```

This function should again accept the output of `planned_analysis()` as input.
The output of this function should be formatted in Markdown.
The idea is to automatically generate the full results section from the analysis.
This way, the preregistration specifies not only the computation but also how the computation's results are reported.
Various packages automatically generate well-formatted Markdown outputs of statistical reports or even entire tables of estimates or figures directly from R objects to help with this objective.
Packages like [`pander`](https://github.com/Rapporter/pander) [@pander], [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf) [@stargazer], [`apaTables`](https://dstanley4.github.io/apaTables/articles/apaTables.html) [@apaTables] and [`papaja`](http://frederikaust.com/papaja_man/) [@papaja] help you to create dynamically generated professional looking results.
The package [`report`](https://github.com/easystats/report) [@report] is particularly noteworthy because it generates not only tables but also a straightforward interpretation of the effects as actual prose (e.g. it verbally quantifies the size of an effect).

Ideally, these three functions can be composed to create a "fake" results section, e.g. when composed to `report_analysis(planned_analysis(simulate_data()))` or `simulate_data() %>% planned_analysis() %>% report_analysis()` outputs a results section.

To summarize, researchers write three functions, `planned_analysis()`, `simulate_data()`, and `report_analysis()` and embed these into a manuscript that serves as a preregistration.
After they gathered the actual data, they replace the simulated data, render the dynamic manuscript (therefore run `planned_analysis()` on the actual data) and write the discussion.

# Preregistration as Code --- a Tutorial

We argued that PAC has several advantages over classic preregistration and have outlined how to implement it.
To illustrate how PAC works in practice and help researchers implementing PAC themselves, we have conducted a PAC that serves as a guiding example.
We will use an exemplary research question that was based on openly available data and merely chosen for simplicity of illustration:

> "Is there a mean difference in the personality trait 'Machiavellism' between self-identified females and males?"

Again, we propose a preregistration format that closely resembles a classic journal article, but uses simulated data and dynamic document generation to create a document that starts out as a preregistration and eventually becomes the final report.
The complete preregistration source is available in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/blob/main/preregistration.Rmd).
In this section, we show code excerpts of this preregistration (formatted in monospace) and explain the rationale behind them.

```{r, echo=FALSE}
show <- function(x, wrap = TRUE, cat = TRUE, fence = TRUE){
  stopifnot(length(x) == 1L)
  if(wrap)x <- str_c(str_wrap(str_split(x, "\\n\\n")[[1]]), collapse = "\n\n")
  if(!isFALSE(fence)){
    if(isTRUE(fence))fence <- "markdown"
    fence <- str_c("````", fence)
    x <-
      str_c(fence,
            x,
            "````",
            sep = "\n",
            collapse = "\n")
  }
  if(cat){
    cat(x)
    return(invisible(x))
  } else {
    return(x)
  }
}
show_prereg <-
  function(section,
           file = "preregistration.Rmd",
           wrap = TRUE,
           cat = TRUE,
           fence = TRUE) {
    prereg <- readr::read_lines(here::here(file))
    start <- stringr::str_c("<!--begin-", section, "-->")
    end <- stringr::str_c("<!--end-", section, "-->")
    start_pos <- which(prereg == start)
    end_pos <- which(prereg == end)
    if (length(start_pos) != 1L || length(end_pos) != 1L)
      stop("No section '", section, "' found.")
    excerpt <-
      str_c(prereg[(start_pos + 1):(end_pos - 1)], collapse = "\n")
    show(excerpt, wrap = wrap, cat = cat, fence = fence)
  }
```

As usual, researchers state why they are interested in their research question in the section "Introduction" and provide the necessary background information and literature to understand the context and purpose of the research question.
This example is drastically shortened for illustration purposes:

```{r, echo=FALSE, results='asis'}
show_prereg("theory")
```

After researchers have provided the research question, they typically proceed to explain how they want to study it.
For simplicity, we will use already published data that we have not yet analyzed:

```{r, echo=FALSE, results='asis'}
show_prereg("method")
```

We choose the following statistical procedure because many researchers are familiar with it[^caution-mww]:

```{r, echo=FALSE, results='asis'}
show_prereg("ttest")
```

[^caution-mww]: The t-test and Mann–Whitney–Wilcoxon test are arguably the most often used hypothesis test [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005 reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005].
The analytical strategy presented here is, in fact, suboptimal concerning several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in the presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]). The interested reader can use the [provided code for the simulation](https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R) to verify that the t-test provides unbiased effect sizes but the Mann–Whitney–Wilcoxon overestimates effect size with increasing sample size and skewness.

The methods section is the translation of the following `planned_analysis()` function:

```{r planned_analysis}
```

This function illustrates two advantages of PAC.
First, a PAC can easily include data-dependent decisions creating different analysis branches under different conditions.
Second, it highlights how difficult it is to describe a statistical analysis precisely.
The same verbal descriptions may be implemented differently by different persons depending on their statistical and programming knowledge and their expectations and beliefs.
One example would be using the function `wilcox.test` instead of the combinations of the functions `rank` and `t.test`.
Either of them are valid implementations of the Mann–Whitney–Wilcoxon test, but the first assumes equal variance.
In contrast, the second applies Welch's correction by default and hence is robust even with unequal variances across groups [@zimmermanRankTransformationsPower1993].
Mentioning every such minute implementation detail is almost impossible and would result in overly verbose preregistrations.
Still, these details can make a difference and provide unwanted researchers' degrees of freedom.

The function `planned_analysis()` together with the function `simulate_data()` (not shown here) can be used to justify the planned sample size.
To that end, `simulate_data()` is repeatedly called with increased sample size and the proportion of significant results (power) is recorded.
The results for such a Monte Carlo simulation for this example are visualized in figure \@ref(fig:power).
The code for this power analysis can be found in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R).
The next snippet shows how we included the results dynamically into the preregistration (the origin of the R-variables `minn`, `choosen_power`, and `choosen_d` is not shown).

```{r, echo=FALSE}
res <-
  read_csv(here::here("data", "simulation_results.csv"), col_types = "ddddddd")
choosen_power <- .8
choosen_d <- .2
minn <- filter(res, d == choosen_d, power > choosen_power) %>% 
  filter(n == min(n))
```

```{r, include=FALSE}
fig_caption <- str_c("Results of simulation for the power analysis. The cross indicates the sample size that archives ", choosen_power*100, "% assuming a Cohen's D of ", choosen_d, ".")
```

```{r power, echo=FALSE, fig.cap=fig_caption}
res %>%
  ggplot(aes(n, power, color = d, group = d)) +
  geom_line() +
  geom_point(
    data = minn,
    size = 3,
    color = "black",
    shape = 3
  ) +
  theme_minimal() +
  scale_color_viridis_c() +
  theme(legend.title = element_text()) +
  labs(y = "Power\n(Proportion of significant results)",
       x = "Sample size",
       color = "Cohen's D") +
  NULL
```

```{r, echo=FALSE, results='asis'}
show_prereg("power")
```

Monte Carlo simulations are, of course, not only applicable for this analysis method and also allow researchers to investigate other relevant properties of their analysis method beyond power [@brandmaier2018precision; @harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].

We implemented a mechanism that only uses simulated data when the actual data is not yet available (in this example if the file `data/data.csv` does not exist) for the results section.
This mechanism also warns readers if these results are based on simulated data.
The warning is colored red to avoid any confusion between mock and actual results.
As soon as the actual data are available, the simulated data is no longer used, and the results represent the actual empirical results of the study.

```{r, echo=FALSE, results='asis'}
show_prereg("results", wrap = FALSE)
```

The last thing we need to do is to report our results with the combination of the functions `planned_analysis()` and `report_analysis()`.

```{r, echo=FALSE, results='asis'}
show_prereg("report", wrap = FALSE)
```

This is an example of how the results could be reported (based on simulated data):

```{r, echo=FALSE, results='asis'}
set.seed(1234)
show(report_analysis(planned_analysis(simulate_data(900, 8, 0.3, 10)), cat = FALSE), fence = "")
```

Of course it is likely that the preregistered code has to be adapted to the actual data.
Changing the preregistration must be discussed.
This is our summary of what we changed:

```{r, echo=FALSE, results='asis'}
show_prereg("discuss-changes")
```

But readers can inspect and judge the changes for themselves on [GitHub](https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b).

The interested reader can find the example rendered as a pdf file with actual data in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/releases/tag/v0.0.2.0-results) and can view the changes we made since preregistering it on this [GitHub page](https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b).

After the researcher is satisfied with their draft preregistration, they should create an archived and unchangeable version of the project that serves as the preregistration.
zenodo.org [@zenodo] is a publicly funded service provider that archives digital artifacts for research and provides digital object identifiers (DOI) for these archives.
While the service is independent of GitHub---in terms of storage facilities and financing---you can link GitHub and zenodo.org.
Please note that you can only link public GitHub repositories to zenodo.
You may log into zenodo.org through your GitHub account.

> Log into zenodo.org → Account → GitHub[^zenodo-github]
[^zenodo-github]: https://zenodo.org/account/settings/github/

After you have linked a GitHub repository you trigger the archivation by creating a GitHub release.
To create GitHub release, navigate to GitHub:

```{r}
usethis::browse_github()
```

And click on Releases → Draft a new release.
Here you can add all relevant binary files but at least a rendered version of the manuscript and the docker image.

We realize that PACs are challenging.
Not every research question or project is suitable for preregistration in general, and many analytical strategies could prove impractical for a PAC.
Therefore, PAC should not be seen as an attempt to invalidate classical preregistration or not-preregistered research but as a complementary approach.
We are convinced that many reasons justify changing the preregistered code, but are equally convinced that readers should judge themselves if they find these deviations from the preregistration meaningful.
PAC enables just that.
Researcher can change their plans and possibly defend their decision, e.g. in the discussion, while readers of the manuscript can clearly see what was changed through version control.

# Discussion

Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019], and therefore this workflow fits well together with other innovations that employ automation, like machine-readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
Automated research projects promise a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].

Central to these innovations is a property we call "reusability" which we hope this workflow promotes.
Reusable code can run on different inputs from a similar context and produce valid outputs.
This property is based on reproducibility but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] such that it is _build-for-reuse_ [@al-badareenReusableSoftwareComponents2010].
The reproducible workflow we present here is heavily automated and hence promotes reusability.
Furthermore, adhering to principles of reusability typically removes errors in the code and thus increases the likelihood that the statistical analysis is correct.
Therefore reproducibility facilitates traditional good scientific practices and provides the foundation for promising innovations.

## Summary

This paper demonstrated how the R package `repro` supports researchers in creating reproducible research projects including reproducible manuscripts.
These are important building blocks for transparent and cumulative science because they enable others to reproduce statistical and computational results and reports later in time and on different computers.
The workflow we present here rests on four software solutions, version control, dynamic document generation, dependency tracking and software management to all but guarantee reproducibility.
We first demonstrated how to create a reproducible research project.
Then, we illustrated how such a project can be reproduced - either by the original author and/or collaborators, or by a third party.
We finally presented an example of how the rigorous and automated reproducibility workflow introduced by `repro` may enable other innovations,
such as the preregistration as code (PAC).
A PAC incorporates all planned analysis code into a reproducible manuscript that already renders the expected results.
These results are based on simulated data, so that there is no danger of peeking at the real data while writing the scripts.
The entire manuscript, including planned analyses, is preregistered.
This way, all researcher's degrees of freedom are eliminated.
Once real data is gathered, the reproducible manuscript is (re-)created with the real data.
Because the same dynamic document serves as preregistration and final manuscript, researcher, reviewers and editors only have to revise one document.
To editors, reviewers and readers it is obvious how the researchers deviated from the preregistration because everything there is a single "source of truth"---dynamic manuscript including code---under version control.

The likelihood of those deviations, while never unavoidable, is decreased by using simulated data to verify the suitability of the proposed analysis plan.
This analysis plan may contain data-dependent decisions to increase flexibility of research plans.

## Limitations

`repro`'s primary goal is to increase the user-friendliness of reproducibility tools for researchers.
Despite our efforts to ensure this goal, the package is still in development and undoubtedly has limitations.
The workflow we present here introduces several software solutions, which in itself present a threat to long-term reproducibility.
To benefit from automatic and convenient reproduction, researcher must have `Git`, `Make`, and `Docker`.
However, `Git` and `Make` are itself included in the `Docker` image created by `repro`.
Researchers can therefore employ the `Docker` image manually to download the `Git` repository and execute `Make` for full reproduction.
In other words the only hard requirement for reproduction and therefore the Achilles' heel for this workflow is `Docker`.
The `Docker` approach has two vulnerabilities.
First, and more important, the `Docker` image for the project and the `Git` repository have to remain available.
The `Dockerfile` (the plain text description to build a `Docker` image) as opposed to the image is insufficient because it relies on too many service providers (e.g. Microsoft R Application Network, Ubuntu Package Archive).
We recommend to archive the `Git` repository and the `Docker` image with zenodo.org, a non-profit long-term storage for scientific data.
The necessary steps for archival on zenodo.org are described in at the end of section [Preregistration as Code --- a Tutorial].
The second vulnerability is that even if the existence of the `Docker` image and `Git` repository is guaranteed, future researchers still require software to run the image.
To that end they can either rely on `Docker` itself or `Docker`-compatible alternatives (e.g. CoreOS rkt, Mesos Containerizer, Singularity).
The only way to remove the reliance on such external software is to turn the `Docker` image into a full operating system that subsequently can be installed and run on almost any modern computer.
This process is technically possible and would guarantee reproducibility for decades without any software dependency, assuming hardware that conforms to the x86 instruction set architecture continues to be available.
However, this process requires much technical knowledge and is currently not facilitated by `repro`.
In any case and without such precautions do the `R Markdown`, the `Makefile` and the `Dockerfile` provide information that allow researchers to manually trace the computational steps and recreate the computational environment.
The `Makefile`, for example, is written in a way that researchers can manually trace the dependencies and execute commands in the right order, in case they are, for whatever reason, unable to run `Make`.
Hypothetically, even if `Docker` were unavailable one day, the `Dockerfile` still serves as unambiguous documentation of how the original system was set up and may help future users to create a software environment that closely resembles the original.

## Outlook

Open science practices are a continually evolving field where technical innovations foster changes in research practice.
Open data is much more widespread thanks to online storage facilities; preregistration are possible because there are preregistration platforms and so forth.
Similarly, we hope that fully automatic reproduction, e.g. with `repro` as technical innovation, increase scientific rigour and efficiency.

This ideal of a fully automatic reproduction of research projects meets in practice a wide range of demands on how user-friendly and how powerful the software solutions are.
Some may find `Make` too complicated or that `Docker` requires too much storage.
Yet others may find that they require other programming languages or scale their computation across hundreds of computers, e.g. via high-performance computing clusters or cloud computing.

`repro` was designed modularly to meet many such demands.
At the moment, `repro` only supports the combination of `R Markdown`, `Git`, `Make`, and `Docker`.
However, there are alternatives for each of these elements that may fit better into an individual research project.
`R Markdown` could be complemented or replaced by a dynamic Microsoft Word document with the help of `officer`[@officer] or `officedown`[@officedown] to accommodate a wider arrange of journal submission standards.
Instead of using formal version control with `Git`, `repro` could automatically save snapshots for increasing user-friendliness.
`Make` could be replaced by the more R-centered alternative `targets` for more convenience.
`Docker` could be combined with `renv` [@R-renv] to control the package versions precisely (our approach fixes the date, `renv` the exact package version).
Alternatively, `Docker` could be replaced by the more lightweight `renv` if no dependencies outside of R are considered crucial.
`Docker` does not satisfy the requirements of many HPC environments, but Singularity was designed to avoid this limitation while still being compatible with `Docker` images.

`repro`'s modular structure allows such alternative workflow, though they have not been implemented.
Depending on the demand by users, we will implement some of these alternative workflows in `repro` and hope for a broad adoption of computational reproducibility in the near future.

# References