### 2021-11-19 manuscript.Rmd
Comments until 1113

---
title: "Reproducible Research in R: A Tutorial on how to Do the Same Thing More Than Once"
author:
  - name: Aaron Peikert
    affil: 1, 2, *
    orcid: 0000-0001-7813-818X
  - name: Caspar J. van Lissa
    affil: 3, 4
    orcid: 0000-0002-0808-5024
  - name: Andreas M. Brandmaier
    affil: 1, 5, 6
    orcid: 0000-0001-8765-6982
affiliation:
  - num: 1
    address: |
      Center for Lifespan Psychology---Max Planck Institute for Human Development, Lentzeallee 94, 14195 Berlin, Germany
  - num: 2
    address: |
      Humboldt-Universität zu Berlin, Berlin, Germany
  - num: 3
    address: |
      Department of Methodology & Statistics---Utrecht University, Faculty of Social and Behavioral Sciences, Utrecht, Netherlands
  - num: 4
    address: |
      Open Science Community Utrecht, Utrecht, Netherlands
  - num: 5
    address: |
      Max Planck UCL Centre for Computational Psychiatry and Ageing Research
      Berlin, Germany, and London, UK
  - num: 6
    address: |
      MSB Medical School Berlin,
      Berlin, Germany
# firstnote to eighthnote
correspondence: |
  peikert@mpib-berlin.mpg.de
journal: psych
type: tutorial
status: submit
bibliography: temp.bib
simplesummary: |
  Reproducibility has long been considered integral to the scientific method.
  An analysis is considered reproducible if an independent person can obtain the same results from the same data.
  Until recently, detailed descriptions of methods and analyses were the primary instrument for ensuring scientific reproducibility.
  Technological advancements now enable scientists to achieve a more comprehensive standard that allows anyone to access a digital research repository and reproduce all computational steps from raw data to final report, including all relevant statistical analyses, with a single command.
  This method has far-reaching implications for scientific archiving,
  reproducibility and replication,
  scientific productivity,
  and the credibility and reliability of scientific knowledge.
  One obstacle to the widespread use of this method is that the underlying tools
  are complex and not part of most researchers' basic training.
  This paper introduces `repro`, an R package that guides researchers through installation and use of the tools required to make a research project reproducible.
  We also suggest using the proposed workflow for the preregistration of study plans as reproducible computer code (preregistration as code; PAC).
  Since computer code represents the planned analyses exactly as they will be executed, it is more precise than natural language descriptions. 
  PAC circumvents the shortcomings of ambiguous preregistrations that may result in undisclosed use of researcher degrees of freedom.
  Reproducibility, facilitated by automation, has a wide range of applications and could potentially accelerate scientific progress.
abstract: "`r tryCatch(trimws(readr::read_file(here::here('abstract.Rmd'))))`"
keywords: |
  open science; computational reproducibility; preregistration; R; R Markdown; Make; GitHub; Docker
acknowledgement: |
  We would like to thank Maximilian Stefan Ernst for his contributions to the code for the simulation study. We are grateful to Julia Delius for her helpful assistance in language and style editing. The R package `repro` was developed as part of the [first author's master thesis](https://aaronpeikert.github.io/repro-thesis/) at the Humboldt-Universität zu Berlin.
authorcontributions: |
  Aaron Peikert took the lead in writing and provided the initial draft of the manuscript.
  Andreas Brandmaier and Caspar J. Van Lissa contributed further ideas, critical feedback, and revisions of the original manuscript.

conflictsofinterest: |
    The authors declare no conflict of interest.
    We have received no financial support for the research, authorship, and/or publication of this article.
abbreviations:
  - short: PAC
    long: Preregistration as Code
  - short: Gb
    long: Gigabyte
  - short: Kb
    long: Kilobyte
  - short: CRAN
    long: Comprehensive R Archive Network
  - short: WORCS
    long: Workflow for Open Reproducible Code in Science
repro:
  packages:
    - tidyverse
    - usethis
    - gert
    - aaronpeikert/repro@15a069
    - here
    - rstudio/webshot2@f62e743
    - targets
    - renv
    - slider
    - patchwork
    - knitr
    - pander
    - lavaan
    - furrr
    - future.batchtools
    - rticles
    - moments
    - report
  scripts:
    - R/simulation.R
    - R/simulation_funs.R
    - R/link.R
  data:
    - data/simulation_results.csv
  images:
    - images/nutshell.pdf
  files:
    - references.bib
  apt:
    - inkscape
output:
  bookdown::markdown_document2:
    base_format: rticles::mdpi_article
header-includes:
   - \DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}, numbers=left}
---

```{r setup, include=FALSE}
library(repro)
automate_load_packages()
automate_load_scripts()
source(here::here("R", "link.R"))
source(here::here("R", "simulation_funs.R"))
```

```{r echo=FALSE}
# define a print method for objects of the class data.frame
knit_print.data.frame = function(x, ...) {
  res = paste(c('', '', kable(x)), collapse = '\n')
  asis_output(res)
}
# register the method
registerS3method("knit_print", "data.frame", knit_print.data.frame)
```

\captionsetup{labelformat=empty}

# Introduction

Scientists increasingly strive to make research data, materials, and analysis code openly available.
Sharing these digital research products can increase scientific efficiency by enabling researchers to learn from each other, reuse materials, and increase scientific reliability by facilitating the review and replication of published research results.
To some extent, these potential benefits are contingent on whether these digital research products are reproducible. 
Reproducibility can be defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code [see @Peikert2019 for details].
The credibility of empirical research results hinges on their objectivity.
Objectivity in this context means, "in principle it [the research finding] can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p. 22].
Only a reproducible result meets these requirements.
Therefore, reproducibility has long been considered integral to empirical research.
Unfortunately, despite increasing commitment to open science practices and good will, many projects can not yet be reproduced by other research teams [@obels2020].
This is because there are various challenges to making a research project reproducible (e.g., missing software dependencies or ambiguous documentation of the exact computational steps taken), and there is a lack of best practices for overcoming these challenges [but see @vanlissa2020worcs].
With technological advancement, however, it is now possible to make all digital products related to a research project available in a manner that enables automatic reproduction of the research project with minimal effort.

With this paper, we pursue two aims.
First, we want to introduce researchers to a notion of automated reproducibility that requires no manual steps, apart from the initial setup of the software environment. 
Secondly, we discuss the implications of automated reproducibility for changing the general approach to research.
With regard to the first goal, we discuss how to address four common threats to reproducibility, using tools originating from software engineering [see @Peikert2019 for details], and present a tutorial on how to employ these tools to achieve automated reproducibility.
A single tutorial cannot comprehensively introduce the reader to the detail of individual tools, but this tutorial is intended to help readers get started with a basic workflow.
The tutorial is aimed at researchers who regularly write code to analyze their data and are willing to make relevant code, data, and materials available, either publicly or on request.
Ideally, the reader has already created a dynamic document at some point in time (e.g., with R Markdown or Jupyter) and used some form of version control (e.g., Git). 
The R package `repro` supports researchers in setting up the required software and in adopting this workflow.
We present automated reproducibility as a best practice; a goal that is not always fully achieved due to limited resources, technical restrictions, or practical considerations, but is worth striving for nonetheless.

In pursuit of the second aim, we present a strictly reproducible and unambiguous form of preregistration [@NosekRevolution2018] that builds upon implementing this reproducible workflow, the so-called *Preregistration as Code* (PAC).
PAC involves preregistering the intended analysis code and the major part of the final scientific report as a dynamic document, including typical sections like introduction, methods, and results.
The resulting dynamic document closely resembles the final manuscript but uses simulated data to generate placeholder results (e.g., figures, tables, and statistics).
Simulated data serve two functions, they allow to test the code for the planned analyses and for preregistering the exact presentation of the results.
Once the empirical data are available, these replace the simulated data; the results are then updated automatically, and the discussion can be written to finalize the report.

Scientific organizations and funding bodies increasingly demand transparent sharing of digital research products, and researchers are increasingly willing to do so.
However, although the sharing of such digital research products is a necessary condition for reproducibility, it is not a sufficient one.
This was illustrated by an attempt to reproduce results from open materials in the journal *Cognition* [@hardwicke2018]. 
Out of 35 published articles with open code and data, results of 22 articles could be reproduced, but in 11 of these cases, further assistance from the original authors was required.
For 13 articles, at least 1 outcome could not be reproduced—--even with the original authors' assistance.
Another study of 62 Registered Reports found that only 41 had data available, and 37 had analysis scripts available [@obels2020].
The authors could execute only 31 of the scripts without error and reproduce the results of only 21 articles (within a reasonable time).
These failed attempts to reproduce findings highlight the need for widely accepted reproducibility standards because open repositories do not routinely provide sufficient information to reproduce relevant computational and statistical results.
If digital research products are available but not reproducible, their added value is limited.

This tutorial demonstrates how R users can make digital research products more reproducible, while striking a balance between rigor and ease-of-use.
A rigorous standard increases the likelihood that a project will remain reproducible as long as possible.
An easy-to-use standard, on the other hand, is more likely to be adopted.
Our approach is to promote broad adoption of such practices by ensuring a "low threshold", by making it easy to get started, while enabling a "high ceiling" by ensuring that they are compatible with more complex rigorous solutions.
As researchers become more proficient in using the tools involved, they can thus further improve the reproducibility of their work.

We have structured the tutorial with a *learning-by-doing* approach in mind, such that readers can follow along on their own computers.
We explicitly encourage readers to try out all R commands for themselves.
Unless stated otherwise, all code blocks are meant to be run in the statistical programming language R (tested with version `r with(R.version, paste0(major, ".", minor))`).

# Threats to Reproducibility and Appropriate Remedies

From our own experience with various research projects, we have identified the following common threats to reproducibility:

1.	*Multiple inconsistent versions of code, data, or both*; for example, the data set may have changed over time because outliers were removed at a later stage or an item was later recoded; or, the analysis code may have been modified during the writing of a paper because a bug was removed at some point in time.
It may then be unclear which version of code and data was used to produce some reported set of results.
2.	*Copy-and-paste errors*; for example, results are often manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result.
3.	*Undocumented or ambiguous order of computation*; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually without documentation (e.g., executing a command manually rather than in a script; copy-and-pasting results from one program to another).
4.	*Ambiguous software dependencies*; for example, a given analysis may depend on a specific version of a specific software package, or rely on software that might not be available on a different computer, or no longer exist at all; or a different version of the same software may produce different results.

We have developed a workflow that achieves long-term and cross-platform computational reproducibility of scientific data analyses.
It leverages established tools and practices from software engineering and rests on four pillars that address the aforementioned causes of non-reproducibility [@Peikert2019]:

1. Version control
2. Dynamic document generation
3. Dependency tracking
4. Software management

The remainder of this section briefly explains why each of these four building blocks is needed and details their role in ensuring reproducibility.
A more extensive treatment of these tools is given in @Peikert2019.

*Version control* prevents the ambiguity that arises when multiple versions of code and data are created in parallel during the lifetime of a research project.
Version control allows a clear link between which results were generated by which version of code and data.
This addresses the first threat to reproducibility, because results can only be said to be reproducible if it is clear which version of data and code produced them.
We recommend using `Git` for version control, because of its widespread adoption in the R community.

`Git` tracks changes to all project-related files (e.g., materials, data, and code) over time.
At any stage, individual files or the entire project can be compared to, or reverted to, an earlier version.
Moreover, contributions (e.g., from collaborators) can be compared to, and incorporated in the main version of the project. 
Version control thus reduces the risk of losing work and facilitates collaboration.
`Git` is built around snapshots that represent the project state at a given point in time.
These snapshots are called *commits* and work like a "save" action.
Ideally, each commit has a message that succinctly describes these changes.
It is good practice to make commits for concrete milestones (e.g., "Commented on Introduction," "Added SES as a covariate," "Address Reviewer 2's comment 3").
This makes it easier to revert specific changes than when multiple milestones are joined in one commit, e.g., "Changes made on 19/07/2021".
Each commit refers back to its ancestor, and all commits are thus linked in a timeline.
The entirety of commits (i.e., the version-controlled project) is called a repository.
In `Git`, specific snapshots of a repository can be tagged, such that the user can clearly label which version of the project was used to create a preregistration, preprint, or final version of the manuscript as accepted by a journal.
`Git` has additional features beyond basic version control, such as "branches" (parallel versions of a project that can later be merged again) to facilitate simultaneous collaboration.
@vuorreCuratingResearchAssets2018 provide a more extensive treatment of how `Git` functions and how to use `Git` for research.
@bryanExcuseMeYou2018 provides additional information on how to track `R Markdown` documents.
Collaborating via `Git` is facilitated by uploading the repository to a cloud-based service.
We recommend GitHub as a host for `Git` repositories because of its popularity among R users.
GitHub has many tools that facilitate working with `Git` --- in particular project management and collaboration --- but these are not central to achieving reproducibility.

Second, we rely on *dynamic document generation*.
The traditional way of writing a scientific report based on a statistical data analysis uses two separate steps conducted in two different programs.
The researcher writes text in a word processor, and conducts the analysis in another program.
Results are then (manually) copied and pasted from one program to another, a process that often produces inconsistencies [@nuijtenPrevalenceStatisticalReporting2016].

Dynamic document generation integrates both steps.
Through dynamic document generation, code becomes an integral, although usually hidden, part of the manuscript, complementing the verbal description and allowing interested readers to gain a deeper understanding of the contents [@knuthCWEBSystemStructured; @claerboutElectronicDocumentsGive1992].
`R Markdown` uses `Markdown` for text formatting and R (or other programming languages) for writing the statistical analysis.
`Markdown` is a lightweight text format in plain text with a minimal set of reserved symbols for formatting instructions.
This way, `Markdown` does not need any specialized software for editing.
It is userfriendly (unlike, for example, `LaTeX` [@lamportLATEXDocumentPreparation1994]), works well with version control systems, and can be exported to various document formats, such as HTML websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
`Markdown` can be used for all sorts of academic documents, ranging from simple sketches of ideas to scientific manuscripts [@R-rticles] and presentations [@revealjs], or even résumés [@vitae].
`R Markdown` extends regular `Markdown` by allowing users to include R code chunks (in fact, arbitrary computer code [@xieMarkdownCookbook2020, Chapter 15, Chapter 15, Other Languages]) into a `Markdown` document.
Upon rendering the document, the code blocks are executed, and their output is dynamically inserted into the document.
This allows the creation of (conditionally) formatted text, statistical results, and figures that are guaranteed to be up-to-date because they are created anew every time the document is rendered to its output format (e.g., presentation slides or a journal article).
@xieMarkdownCookbook2020 provides an extensive yet practical introduction to most features of `R Markdown`.

While version control and dynamic document generation are becoming more common, we have argued that two more components are required and that each component alone is unlikely to guarantee reproducibility [@Peikert2019; @vanlissa2020worcs].
In practice, dependencies between project files (e.g., information on what script uses which data file and what script needs to be run first) or on external software (e.g., system libraries or components of the programming language, such as other R packages) are frequently unmentioned or not exhaustively and unambiguously documented.

*Dependency tracking* helps automatically resolve dependencies between project files.
In essence, researchers provide a collection of *computational recipes*.
A computational recipe describes how inputs are processed to deterministically create a specific output in a way that is automatically executable.
The concept of computational recipes is central to our understanding of reproducibility because it enables a unified way to reproduce a project automatically.
Similar to a collection of cooking recipes, we can have multiple products (_targets_) with different ingredients (_requirements_) and different steps of preparation (_recipes_).
In the context of scientific data analysis, targets are typically the final scientific report (e.g., the one to be submitted to a journal) and possibly intermediate results (such as preprocessed data files, simulation results, and analysis results).
A workflow that involves renaming variable names by hand in a graphical spreadsheet application, for example, is therefore incompatible with automated reproducibility.
Another property of a computational recipe is that the same inputs should always result in the same outputs.
For most computer code (given the same software is used), this property is fulfilled.
However, one noteworthy exception is the generation of pseudo-random numbers.
Whenever random numbers are used in a computation, it is only reproducible if the random number generator generates the same numbers.
To ensure identical random numbers, users may fix the state of the random number generated with a so-called seed (e.g. `set.seed()` in R), but they also need to guarantee that the pseudo-random number generator is unchanged [see @Peikert2019].

We recommend using `Make` for dependency tracking because it is language independent.
The following hypothetical example illustrates the utility of `Make` and a suitable `Makefile`.
Consider a research project that contains a script to simulate data (`simulate.R`) and a scientific report of the simulation results written in `R Markdown` (`manuscript.Rmd`).
A `Makefile` for this project could look like this:

```{bash, eval=FALSE}
manuscript.pdf: manuscript.Rmd simulated_data.csv
  Rscript -e 'rmarkdown::render("manuscript.Rmd")'

simulated_data.csv: simulate.R
  Rscript -e 'source("simulate.R")'
```

There are two targets, the final rendered report (`manuscript.pdf`, l. 1) and the simulation results (`simulation_results.csv`, l. 4).
Each target is followed by a colon and a list of requirements.
If a requirement is newer than the target, the recipe will be executed to rebuild the target.
If a requirement does not exist, `Make` uses a recipe to build the requirement before building the target.
Here, if one were to build the final `manuscript.pdf` by rendering the `R Markdown` with the command shown in l.2, `Make` would check whether the file `simulation_results.csv` exists; if not, it would issue the command in l.5 to run the simulation before rendering the manuscript.
This ensures that the simulated data are present before the manuscript is built, and that the simulation is re-run and the manuscript is rebuilt if the simulation code was changed.
`Make` therefore offers a standardized process to reproduce projects, regardless of the complexity or configuration of the project.
Note that the Workflow for Open Reproducible Code in Science (WORCS) we presented elsewhere [@vanlissa2020worcs] does not explicitly contain this dependency tracking element, but its strict structure of only containing one definite `R Markdown` still makes dependencies between files unambiguous.

```{r, include=FALSE}
container_size <- as.numeric(fs::file_size("reprotutorial.sif"))/1024^3
```

A version-controlled dynamic document with dependency tracking still relies on external software.
Troubleshooting issues specific to a particular programming language or dependent tool typically require considerable expertise and threaten reproducibility.
*Software management* refers to the act of providing records of, or access to, all software packages and system libraries a project depends on.
One comprehensive approach to software management is containerization.
The central idea is that by
"[..] packaging the key elements of the computational environment needed to run the desired software [makes] the software much easier to use, and the results easier to reproduce [...]" [@silverSoftwareSimplified2017, p. 174].

*`Docker`* is a popular tool for containerization.
It manages software dependencies by constructing a virtual software environment independent of the host software environment.
These so-called "`Docker` images" function like a virtual computer (i.e., a "sand box" a computational environment seperated from the host).
A `Docker` image contains *all* software dependencies used in an analysis---not just R packages, but also R and Rstudio, and even the operating system.
This is important because low level functionality may impact the workings of higher-order software like R, such as calls to random number generators or linear algebra libraries.
All of the differences in computational results that could be caused by variation in the software used are hence eliminated.

Note that the software environment of the Docker image is completely separate from the software installed on your computer.
This separation is excellent for reproducibility but takes some getting used to.
For example, it is important to realize that software available on your local computer will *not* be accessible within the confines of the `Docker` image.
Each dependency that you want to use within the `Docker` image must be explicitly added as a dependency.
Furthermore, using `Docker` may require you to install software on an operating system that may not be familiar to you.
The images supplied by @boettigerIntroductionRockerDocker2017, for example, are based on Linux.

There are two ways to build a `Docker` image.
First, users can manually install whatever software they like from within the virtual environment.
Such a manually build environment can still be ported to all computers that support `Docker`.
However, we prefer the second way of building images automatically from a textual description called `Dockerfile`.
Because the `Dockerfile` clearly describes how which software is installed, the installation process can be repeated automatically.
Users can therefore quickly change the software environment, for example, update to another R version or given package version. 
Packaging all required software in such an image requires considerable amounts of storage space.
Two major strategies help to keep the storage requirements reasonable.
One is to rely on pre-made images that are maintained by a community for particular purposes.
For example, there are pre-made images that only include what is necessary for R, based on Ubuntu containers [@boettigerIntroductionRockerDocker2017].
Users can then install whatever they need in addition to what is provided by these pre-compiled images.
`r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}GiB of disk space."))`
The image for this project includes Ubuntu, R, RStudio, LaTeX as well as a variety of R packages like tidyverse [@tidyverse] and all its dependent packages, amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` R packages.

A second strategy is to save a so-called  `Dockerfile`, which contains only a textual description of all commands that need to be executed to recreate the software environment.
`Dockerfiles` are tiny (the [`Dockerfile`](https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile) for this project has a size of only `r round(as.numeric(fs::file_size("Dockerfile"))/1024, 2)`KiB).
However, they rely on the assumption that all software repositories that provide the dependent operating systems, pieces of software, and R packages will continue to remain accessible and provide historic software versions.
For proper archiving, we therefore recommend storing a complete image of the software environment, in addition to the `Dockerfile`.
A more comprehensive overview of the use of containerization in research projects is given by @wiebelsLeveragingContainersReproducible2021.
Note that WORCS, which we presented elsewhere [@vanlissa2020worcs] relies on the R package `renv` [@R-renv] for software management.
`renv` is more lightweight and easier to use than `Docker` on the one hand, but not as comprehensive on the other because it only takes snapshots of the R packages instead of all software used.

To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)) dynamic document generation (using `R Markdown`), version control (using `Git`), dependency tracking (using `Make`), and software management (using `Docker`).
While `R Markdown` and `Git` are well integrated into the R environment through RStudio, `Make` and `Docker` require a level of expertise that is often beyond the training of scholars outside the field of information technology.
This presents a considerable obstacle to the acceptance and implementation of the workflow.
To overcome this obstacle, we have developed the R package `repro` that supports scholars in setting up, maintaining, and reproducing research projects in R.
Importantly, a reproducible research project created with `repro` does not have the `repro` package itself as a dependency.
These projects will remain reproducible irrespective of whether `repro` remains accessible in future.
Users do not need to have `repro` installed to reproduce a project;
in fact, they do not even need to have R installed because the entire project can be rebuilt inside a container with R installed.
In the remainder, we will walk you through the creation of a reproducible research project with the package `repro`.

```{r schematic, eval = TRUE, echo = FALSE, fig.cap="Schematic illustration of the interplay of the four components (in dashed columns) central to the reproducible workflow: version control (`Git`), dependency tracking (`Make`), software management (`Docker`), and dynamic document generation (`R Markdown`). `Git` tracks changes to the project over time. `Make` manages dependencies among the files. `Docker` provides a container in which the final report is built using dynamic document generation in `R Markdown`. Adapted from @Peikert2019 licensed under [CC BY 4.0](https://creativecommons.org/licenses/by/4.0).  "}
# file gets downloaded in Makefile
knitr::include_graphics(here::here("images/nutshell.svg"), auto_pdf = TRUE)
```

# Creating Reproducible Research Projects

One impediment to the widespread adoption of a standard for reproducible research is that learning to use the required tools can be quite time-intensive.
To lower the threshold, the R package `repro` introduces helper functions that simplify the use of complicated and powerful tools.
The `repro` package follows the format of the [`usethis`](https://usethis.r-lib.org) [@usethis] package, which provides helper functions to simplify the development of R packages.
The `repro` package provides similar helper functions, but focuses on reproducibility-specific utilities.
These helper functions guide end-users in the use of reproducibility tools, provide feedback about what the computer is doing and suggest what the user should do next.
We hope this makes reproducibility tools more accessible by enabling beginner-level users to detect their system's state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
These wrappers are merely a support system; as users learn to use the underlying tools, they can rely less on `repro` and use these ttools directly to solve more complex problems.

This tutorial assumes that the user will be working predominantly in R, with the help of RStudio.
It describes basic steps that we expect to be relevant for small-scale psychological research projects that do not rely on external software or multistage data processing (for those requirements see section [Advanced Features]). 
Of course, your specific situation might involve additional, more specialized steps. 
After completing the tutorial, you should be able to customize your workflow accordingly.

The first step is to install the required software.
We assume that you have installed R [@R-base, version `r with(R.version, paste0(major, ".", minor))`] and RStudio [@rstudio, version 1.4] already but the tutorial will guide you in detail through the installation of other necessary software with the help of the R package [`repro`](https://github.com/aaronpeikert/repro) [@R-repro].
In case you have either not installed R and RStudio or are unsure if they are up-to-date, you might want to consult our installation advice in the [Online Supplementary Material](https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md) that covers the installation of all software necessary for this tutorial in three steps.
The [installation advice]((https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md)) may also help Windows users who have problems installing `Docker`.

Unfortunately, Docker requires administrator rights to run, which may not be available to all researchers.
We recommend [`renv`](https://rstudio.github.io/renv/articles/renv.html) [@R-renv] in cases where no administrator rights can be obtained but can not detail its use in this document.
`renv` tracks which R package is installed from which source in which version in a so-called [`lockfile`](https://rstudio.github.io/renv/articles/lockfile.html).
This lockfile is then used to reinstall the same packages on other computers or later in time.
For a more thorough discussion, see @vanlissa2020worcs.

Start RStudio and install the package [`repro`](https://github.com/aaronpeikert)[@R-repro]. 
It will assist you while you follow the tutorial.

```{r, eval=FALSE}
# repro is not on CRAN yet
options(
  repos = c(aaronpeikert = 'https://aaronpeikert.r-universe.dev',
            CRAN = 'https://cloud.r-project.org')
)
install.packages('repro')
```

To verify that you have indeed installed and set up the required software for this workflow,
you can use the "check functions".
These also illlustrate how `repro` assists the user in setting up a reproducible workflow.
In the example below, we use the double-colon operator to explicitly indicate which functions originate in the `repro` package.
If the package is loaded (using `library("repro")`), it is not necessary to use this double-colon notation.

```{r, include=FALSE}
opts <- options()
options(repro.docker = TRUE,
        repro.docker.running = FALSE)
```

```{r}
# `package::function()` → use function from package without `library(package)`
repro::check_git()
repro::check_make()
repro::check_docker()
```

```{r, include=FALSE}
options(opts)
```

These functions check whether specific dependencies are available on the user's system, and if not, explain what further action is needed to obtain it.
Sometimes they ask the user to take action; for example, the following happens if you are a Windows user who does not have `Git` installed:

```{r, include=FALSE}
opts <- options()
options(repro.os = "windows",
        repro.git = FALSE)
```

```{r}
repro::check_git()
```

```{r, include=FALSE}
options(opts)
```

The messages from `repro` try to help the user solve problems.
They are adjusted to your specific operating system and installed dependencies.
Before you continue, we ask you to run the above commands to check `Git`, `Make`, and `Docker`---both to become familiar with the functionality of the `check_*()` functions and to make sure your system is prepared for the remainder of this tutorial.

After you have installed the necessary software, we suggest that you set up a secure connection to GitHub:

```{r, include=FALSE}
opts <- options()
options(repro.github = TRUE,
        repro.github.ssh = TRUE,
        repro.ssh = TRUE,
        repro.github.token = TRUE)
```

```{r}
repro::check_github()
```

If you know what Secure Shell (SSH) is and want to use it, you may alternatively use:

```{r}
# only an alternative: DO NOT USE if you are unsure what SSH means
repro::check_github(auth_method = "ssh")
```

```{r, include=FALSE}
options(opts)
```

If necessary, follow any instructions presented until all checks are passed.

## Creating an RStudio Project

We start by creating a project folder with RStudio by clicking the menu item:

> File → New Project... → New Directory → Example Repro Template

This creates a project with a sample analysis.
This sample analysis consists of a single R Markdown document and a single data file.
The only special thing about the R Markdown document is the `repro` metadata that we will learn later about. 
However, you may turn any other template or existing R project into a reproducible research project by adding those `repro` metadata there.

## Implementing Version Control

Now that your project is set up, we will introduce you to version control with `Git`.
`Git` does not automatically track all files in your project folder;
rather, you must manually add files to the `Git` repository.
To make sure you do not accidentally add files that you do not wish to share (e.g., privacy-sensitive data),
you can list specific files that you do not want to track in the `.gitignore` file.
You can also block specific filetypes; for example, to prevent accidentally sharing raw data.
You can add something to the `.gitignore` file directly or with this command:

```{r, eval=FALSE}
usethis::use_git_ignore("private.md")
```

Now the file `private.md` will not be added to the `Git` repository, and hence also not be made public if you push the repository to a remote service like GitHub.
Please also consider carefully whether you can include data in the repository without violating privacy rights.
If you are not allowed to share your data publicly, add the data file(s) to the `.gitignore` file and only share them on request.

New users are advised to explicitly exclude any sensitive files before proceeding.
When you are ready, you can begin tracking your remaining files using `Git` by running:

```{r, eval=FALSE}
usethis::use_git()
```

```{r modified, eval = TRUE, echo = FALSE, fig.cap="The Git pane in R Studio, showing `manuscript.Rmd` modified but unstaged and `modified.png` newly added and staged.  "}
knitr::include_graphics(here::here("images", "modified.png"), auto_pdf = TRUE)
```

For `Git` to recognize changes to a given file, you have to stage and then commit these changes (this is the basic save action for a project snapshot).
One way to do this is through the visual user interface in the RStudio `Git` pane (see Figure \@ref(fig:modified)).
Click on the empty box next to the file you want to stage.
A checkmark then indicates that the file is staged.
After you have staged all of the files you want, click on the commit button, explain in the commit message why you made those changes, and then click on commit.
This stores a snapshot of the current state of the project.

The files you created and the changes you made have not yet left your computer.
All snapshots are stored in a local repository in your project folder.
To back up and/or share the files online, you can push your local repository to a remote repository.
While you can choose any `Git` service (like GitLab or BitBucket), we will use GitHub in this tutorial.
Before you upload your project to GitHub, you need to decide whether you would like the project to be publicly accessible (viewable by anyone, editable by selected collaborators) or if you want to keep it private (only viewable and editable by selected collaborators).
To upload the project publicly to GitHub use:

```{r, eval=FALSE}
usethis::use_github()
```

To upload it privately:

```{r, eval=FALSE}
usethis::use_github(private = TRUE)
```

Depending on your computer's configuration, it may ask you to set up a secure connection to GitHub.
In this case, first, follow the suggestions shown on the R console.

## Using Dynamic Document Generation

Now that you have created a version-controlled project, we will proceed with dynamic document generation.
A dynamic document has three elements:

1. Text (prose; e.g., a scientific paper or presentation)
2. Executable code (e.g., analyses)
3. Metadata (e.g. title, authors, document format)

`R Markdown` is a type of dynamic document well-suited to the RStudio user interface.
The text of an `R Markdown` is formatted by `Markdown` (see [@xieMarkdownDefinitiveGuide2019] for technical details and [@xieMarkdownCookbook2020] for practical guidance).
The code mostly consists of R code (although other programming languages are supported, like Python, C++, Fortran, etc).
The following example serves to illustrate the `Markdown` syntax. 
It shows how to create a heading, a word in bold font, a citation, and a list of several items in `Markdown`:

```markdown
<!--this is a Markdown file -->
# Heading (level 1)

Normal text.
Important **word** in bold.
A citation: @einstein1935 did important research on this topic.

## Subheading (level 2)

To do list:

* Do research
* Do more research
* Spend time with family and friends
```

One advantage of this type of markup for formatting is that it can be rendered to many different output formats---both in terms of file types, like `.docx`, `.html`, `.pdf`, and in terms of style, e.g. specific journal requirements.
For social scientists, the [`papaja`](https://github.com/crsh/papaja) package [@papaja] may be relevant, as it produces manuscripts that follow the American Psychological Association formatting requirements [@apa7].
`R Markdown` files are plain text, which is more suitable for version control using `Git` than binary files generated by some word processors.
Some users might find it easier to activate the "Visual Editor" of RStudio ([Ctrl] + [Shift] + [F4] or click on the icon that resembles drawing materials or a compass in the upper right corner of the `R Markdown` document), which features more graphical elements like a traditional word processor but still creates an `R Markdown` underneath with all of its flexibility.
The visual editor has some additional benefits, such as promoting best practices (for example, each sentence should be written on a new line, which makes it easier to track changes across versions) and improving the generation of citations and references to tables and figures.

Now that you are familiar with `Markdown` formatting basics, we turn our attention to including code and its results in the text.
Code is separated by three backticks (and the programming language in curly brackets) like this:

````markdown
This is normal text, written in Markdown.

`r ''````{r}
# this is R code
1 + 1
```
````

The hotkey [Control]+[Alt]+[i] inserts a block of code in the file.
The results of code enclosed in such backticks will be dynamically inserted into the document (depending on specific settings).
This means that whenever you render the `R Markdown` to its intended output format, the code will be executed and the results updated.
The resulting output document will be static, e.g., a pdf document, and can be shared wherever you like, e.g., on a preprint server.

Once the `R Markdown` file has been rendered to a static document, the resulting file is decoupled from the `R Markdown` and the code that created it.
This introduces a risk that multiple versions of the static document are disseminated, each with slightly different results.
To avoid ambiguity, we, therefore, recommend referencing the identifier of the `Git` commit at the time of rendering in the static document.
Simply put, a static document should link to the version of the code that was used to create it.
The `repro` package comes with the function `repro::current_hash()` for this purpose.
This document was created from the commit with the hash [``r repro::current_hash()`` (view on GitHub)](`r stringr::str_c("https://github.com/aaronpeikert/repro-tutorial/tree/", repro::current_hash(40))`).

Now that you know how to write text and R code in an `R Markdown`, you need to know about metadata (also called: YAML front matter).
These metadata contain information about the document, like the title and the output format.
Metadata are placed at the beginning of the document and are separated from the document body by three dashes.
The following example is a full markdown document where the metadata (the "YAML front matter") are in lines 1--6.
Some metadata fields are self-explanatory (like the author field), and exist across all output formats (like the title field).
Others are specific to certain output formats or R packages.

````markdown
---
title: "A Tutorial on how to Do the Same Thing More Than Once"
author: Aaron Peikert, Caspar J. van Lissa, and Andreas M. Brandmaier
abstract: A hitchhiker's guide to reproducible research in R
output: html_document
---

# Introduction

Important for reproducibility:

1. *Version control*
2. *Dynamic document creation*
3. *Dependency tracking*
4. *Software management*

`r ''````{r}
# this is R code
t.test(extra ~ group, data = sleep)
```
````

## Manage Software and File Dependencies

The `repro` package adds fields to the metadata to list all dependencies of the research project.
This includes R scripts, data files, and external packages.
The format is as follows (see everything below the line *repro:*):

```markdown
---
title: "A tutorial on how to do the same thing more than once"
author: Aaron Peikert, Caspar J. Van Lissa and Andreas M. Brandmaier
output: html_document
repro:
  scripts:
    - R/load.R
  data:
    - data/mtcars.csv
  packages:
    - tidyverse
    - usethis
    - gert
---
```

This information clarifies what dependencies (in the form of files and R packages) a project relies on.
`repro` uses this information to construct a `Makefile` for the dependencies on other files and a `Dockerfile` that includes all required packages.
Together, these two files form the basis for consistency within a research project and consistency across different systems.
The function `repro::automate()` converts the metadata from all `R Markdown` files in the project (all files with the ending `.Rmd`) to a `Makefile` and a `Dockerfile`.
These files allow users (including your future self) to reproduce every step in the analysis automatically.
Please run `repro::automate()` in your project:

```{r, eval=FALSE}
repro::automate()
```

It is important to re-run `repro::automate()` whenever you change the `repro` metadata, change the output format, or add a new `R Markdown` file to the project to keep the `Makefile` and `Dockerfile` up to date.
There is no harm in running it too often.
Other than the `Makefile` and the `Dockerfile`, which are created in the document root path, `repro` generates a few more files in the `.repro` directory (which we will explain in detail later), all of which you should add and commit to `Git`.

## Reproducing a project

If someone (including you) wants to reproduce your project, they first have to install the required software, that is `Make`, and `Docker`.
Remember, you can use the `check_*`-functions to test if these are installed:

```{r, include=FALSE}
opts <- options()
options(repro.docker = TRUE,
        repro.docker.running = FALSE)
```

```{r}
repro::check_make()
repro::check_docker()
```

```{r, include=FALSE}
options(opts)
```

When these are set up, they can ask `repro` to explain how they should use `Make` and `Docker` to reproduce the project (or you could explain it to them):

```{r}
repro::reproduce()
```

If you feel uncomfortable using the terminal directly, you can send the command to the terminal from within R:

```{r, eval=FALSE}
system(repro::reproduce())
```

The only "hard" software requirement for reproducing a project is `Docker`, assuming users know how to build a `Docker` image and run `Make` within the container.
However, if they have installed `Make` in addition to `Docker`, they do not even need to know how to use `Docker` and can simply rely on the two `Make` commands "`make docker`" and "`make -B DOCKER=TRUE`".

## Summary

1. Install the `repro` package:

```{r, eval=FALSE}
options(
  repos = c(aaronpeikert = 'https://aaronpeikert.r-universe.dev',
            CRAN = 'https://cloud.r-project.org')
)
install.packages('repro')
```

2. Check the required software:

```{r, include=FALSE}
opts <- options()
options(repro.github = TRUE,
        repro.github.ssh = TRUE,
        repro.github.token = TRUE)
```

```{r}
repro::check_git()
repro::check_github()
repro::check_make()
repro::check_docker()
```

```{r, include=FALSE}
options(opts)
```

3. Create an R project or use an existing one. Do not forget to add `repro` metadata (i.e., packages, scripts, data).

```yaml
repro:
  scripts:
    - R/load.R
  data:
    - data/mtcars.csv
  packages:
    - tidyverse
```

The sample `repro` project already has theese metadata:

```{r,eval=FALSE}
repro::use_repro_template("/some/folder")
```

4. Let `repro` generate `Docker`- and `Makefile`:

```{r, eval=FALSE}
repro::automate()
```

5. Enjoy automated reproducibility:

```{r}
repro::reproduce()
```

# Advanced Features

This section is for advanced users who want to overcome some limitations of `repro`.
If you read this paper the first time, you will probably want to skip this section and continue reading from the section "Preregistration as Code."
As explained above, `repro` is merely a simplified interface to the tools that enable reproducibility.
This simplified interface imposes two restrictions.
Users who ask themselves either, "How can I install software dependencies outside of R in the `Docker` image?" or "How can I express complex dependencies between files (e.g., hundreds of data files are preprocessed and combined)?" need to be aware of these restrictions and require a deeper understanding of the inner workings of `repro`.
Other users may safely skip this section or return to it if they encounter such challenges.

The first restriction is that users must rely on software that is either already provided by the base `Dockerimage` "rocker/verse" or the R packages they list in the metadata.
The metadata the `repro::automate()` function relies on can only express R packages as dependencies for the `Dockerfile` and only trivial dependencies (in the form of "file must exist") for the `Makefile`. 
Other software that users might need, like other programming languages, not yet installed LaTeX packages, etc., must be added manually.
We plan to add support for commonly used ways to install software beyond R packages via the metadata and `repro::automate()`, for example, for system libraries (via `apt` the Ubuntu package manager), LaTeX packages (via `tlmgr` the Tex Live package manager), Python packages (via `pip` the python package manager).
The second limitation is related to dependencies.
`Make` can represent complex dependencies, for example: A depends on B, which in turn depends on C and D.
If B is missing in this example, `Make` would know how to recreate it from C and D.
These dependencies, and how they should be resolved, are difficult to represent in the metadata.
Users, therefore, have to either "flatten" the dependency structure by simply stating that A depends on B, C, and D, thereby leaving out important information or express the dependencies directly within the `Makefile`.

The following section explains how to overcome these limitations despite reliance on the automation afforded by `repro`.
Lifting these restrictions requires the user to interact more directly with `Make` or `Docker`.
Users need to understand how repro utilizes `Make` and `Docker` internally to satisfy more complicated requirements.

Let us have a closer look at the command for reproducing a repro project: `make docker && make -B DOCKER=TRUE`; which consists of two processing steps.
First, it recreates the virtual software environment (`Docker`), and then it executes computational recipes in the virtual software environment (`Make`).
The first step is done by the command `make docker`.
The command `make docker` will trigger `Make` to build the target called `docker`.
The recipe for this target builds an image from the `Dockerfile` in the repository.
The `&&` concatenates both commands and only runs the second command if the first is successful.
Therefore, the computational steps are only executed when the software environment is set up.
The second step executes the actual reproduction and is again a call to `Make` in the form of `make -B DOCKER=TRUE` with three noteworthy parts.
First, a call to `make` without any explicit target will build the `Make` target `all`.
Second, the flag `-B` means that `Make` will consider all dependencies as outdated and will hence rebuild everything.
Third, repro constructs `Make` targets so that if you supply `DOCKER=TRUE` they are executed within the `Docker` image of the project.

The interplay between `Docker` and `Make` resembles a chicken or egg problem.
We have computational steps (`Make`) that depend on the software environment (`Docker`) for which we again have computational steps that create it.
Users only require a deeper understanding of this interdependence when they either want to have more complex computational recipes than rendering an `R Markdown` or require other software than R packages.

Users can have full control over the software installed within the image of the project.
`repro` creates three `Dockerfiles` inside the `.repro` directory.
Two `Dockerfiles` are automatically generated.
The first is `.repro/Dockerfile_base`.
It contains information about the base image on which all the remaining software is installed.
By default we rely on the "verse" images provided by the Rocker project [@boettigerIntroductionRockerDocker2017].
These contain (among other software) the packages `tidyverse`, `rmarkdown`, and a complete LaTeX installation, which makes these images ideal for the creation of scientific manuscripts.
Users can choose which R version they want to have inside the container by changing the version number in line 1 to the desired R version number.
By default, the R version corresponds to the locally installed version on which `repro::automate()` was called the first time.
The build date is used to install packages in the version that was available on the Comprehensive R Archive Network on this specific date and can also be changed.
By default, this date is set to the date on which `repro::automate()` was called the first time. 
This way, the call to the automate function virtually freezes the R environment to the state it was called the first time inside the container. 
Below, you see the `Docker` base file we used to create this manuscript:

```{r, results='asis', echo=FALSE}
cat(
  "```bash",
  readLines(here::here(".repro", "Dockerfile_base"))[1:3],
  "```",
  sep = "\n"
)
```

The second automatically generated `Dockerfile` is `.repro/Dockerfile_packages`.
Whenever `repro::automate()` is called, `repro` gathers all R packages from all `.Rmd` files and determines whether they should be installed from CRAN or GitHub fixed to the date specified in `Dockerfile_base`.
Finally, there is one manually edited `Dockerfile`: `.repro/Dockerfile_manual`.
It is blank by default and can be used to add further dependencies outside of R, like system libraries or external software.
Using `Docker` may require you to install software on an operating system that may not be familiar to you.
The images supplied by [@boettigerIntroductionRockerDocker2017], for example, are based on the Ubuntu operating system.
The most convenient way to install software on Ubuntu is through its package manager `apt`.
If the following snippet is added to `.repro/Dockerfile_manual`, the `Docker` image will have, for example, Python installed.
Other software is installed identically, only the software name is exchanged.

```bash
RUN apt-get update && apt-get install -y python3
```

`Docker` eventually requires a single `Dockerfile` to run, so `repro::automate()` simply concatenates the three `Dockerfiles` and saves the result into the main `Dockerfile` at the top level of the R project.
With this approach, users of `repro` can build complex software environments and implement complex file dependencies.
The standard `repro` metadata only make sure that all dependencies are available but does not allow you to specify custom recipes for them in the metadata.
If you can formulate the creation of dependencies in terms of computational steps, e.g. the file `data/clean.csv` is created from `data/raw.csv` by script `R/preprocess.R`, you should include these in the `Makefile`.
The `Makefile` that `repro` creates is only a template, and you are free to change it.
However, make sure you never remove the following two lines:

```bash
include .repro/Makefile_Rmds
include .repro/Makefile_Docker
```

The file `.repro/Makefile_Rmds` contains the automatically generated targets from `repro::automate()` for the `R Markdown` files.
This file should not be altered manually.
If you are not satisfied with the automatically generated target, simply provide an alternative target in the main `Makefile`.
Targets in the main `Makefile` take precedent.

The file `.repro/Makefile_Docker` does again contain a rather complicated template that you could, but should usually not modify.
This `Makefile` coordinates the interplay between `Make` and `Docker` and contains targets for building (with `make docker`) and saving (with `make save-docker`) the `Docker` image.
Additionally, it provides facilities to execute commands within the container.
If you write a computational recipe for a target, it will be evaluated using the locally installed software by default.
To evaluate commands inside the `Docker` image instead, you should wrap them in `$(RUN1) command $(RUN2)`, as done in this example, which is identical to the first `Make` example we gave above except for the addition of `$(RUN1)` and `$(RUN2)` in l. 2:

```bash
simulated_data.csv: R/simulate.R
  $(RUN1) Rscript -e 'source("R/simulate.R")' $(RUN2)
```

If users execute this in the terminal:

```bash
make data/simulation_results.csv
```
It behaves exactly as in the first `Make` example, the script `R/simulate.R` is run using the locally installed R.
Because this translates simply to:

```{r, echo=FALSE, results='asis'}
cat("```bash",
  stringi::stri_wrap(system2("make", " --dry-run data/simulation_results.csv", stdout = TRUE)),
  "```", sep = "\n")
```

But if users use

```bash
make DOCKER=TRUE data/simulation_results.csv
```

it is evaluated within the `Docker` container using the software within it and not the locally installed R version:

```{r, echo=FALSE, results='asis'}
cat("```bash",
  stringi::stri_wrap(system2("make", " --dry-run DOCKER=TRUE simulated_data.csv", stdout = TRUE)),
  "```", sep = "\n")
```

To summarize, `repro` automates dependency tracking (in the form of `Make`) and software management (using `Docker`) without the necessity to learn both tools,
but users with advanced requirements can still customize all aspects of both programs.

# Preregistration as Code

*Preregistration* refers to the practice of defining research questions and planning data analysis before observing the research outcomes [@NosekRevolution2018].
It serves to separate a-priori planned and theory-driven (confirmatory) analyses from unplanned and post-hoc (exploratory) analyses.<!-- CJ I removed the bit about credibility, because that requiresfurther explanation of why separating these two types results in more credible findings-->
Exploratory analyses refer to both those aimed at discovering patterns in data, and those aimed at overcoming shortcomings of confirmatory analyses that come to light after seeing the data (e.g., violations of assumptions, modifications to address poor model fit).
Researchers are faced with a myriad of choices in designing, executing, and analyzing a study, often called researchers degrees of freedom.
Undisclosed researcher degrees may be used to modify planned analyses until a key finding reaches statistical significance or to inflate effect size estimates, a phenomenon referred to as *opportunistic bias* [@decosterOpportunisticBiasesTheir2015].
Preregistration increases transparency by clarifying when and how researchers employ their degrees of freedom.
It expressly does not restrict what researchers may do to gather or analyze their data.

There are still several shortcomings to preregistration.
One is that written study plans are often multiinterpretable.
Empirical research has shown that, even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis, their results vary considerably [@silberzahnManyAnalystsOne2018].
The current best practice to ensure comprehensive and specific preregistration is to impose structure by following preregistration templates [@bowmanOSFPreregTemplate2020, @bakkerEnsuringQualitySpecificity2020].
However, such templates cannot ensure full transparency because it is impossible to verbally describe every detail of an analysis for any but the most straightforward analysis.
This ambiguity causes a second problem, namely, comparing the initial plan and the resulting publication to decide if and how researchers deviated from the preregistration.
This task is difficult because it is impossible to decide without additional information weather the analysis was actually carried out differently or just described differently.
Even if researchers were faithful to the preregistration, readers may reach opposite conclusions because they have to compare two different text that may be worded differently or describe the same thing in varying levels of detail.
A third limitation is that preregistrations are susceptible to non-reproducibility, just like primary research.
To illustrate, a review of 210 preregistrations found that, even though 174 (67%) included a formal power analysis, only 34 (20%) of these could be reproduced [@bakkerRecommendationsPreregistrationsInternal2020].
Even when researchers have gone to great lengths in preregistering an analysis script,
they sometimes inexplicably fail to reproduce their own results.
For example, @steegenMeasuringCrowdAgain2014 realized after publication that part of their preregistered code resulted in different test statistics than they reported initially (see their Footnote 7).
A final limitation is that written plans may turn out to be unfeasible once data are obtained and analyzed.
For example, a verbal description of a statistical model may be unidentified, e.g., if it includes reciprocal paths between variables or more parameters than observed data.
Conversely, a model may be misspecified in a major way; for example, by omitting direct effects when the research question is about mediation, thus leading to a model with an unacceptable fit.
Many researchers would only realize that such a model cannot be estimated once the data are obtained, thus necessitating a deviation from the preregistered plans.

The workflow described in this paper facilitates a rigorous solution to this problem:
Instead of describing the analysis in prose, researchers include the code required to conduct the analysis in the preregistration.
We term this approach of writing and publishing code at the preregistration stage *Preregistration as Code (PAC)*.
PAC has the potential to eliminate undisclosed researchers degrees of freedom to a much greater extent than, e.g., preregistration templates.
Moreover, it reduces overhead by obviating the need to write a separate preregistration and manuscript.
For PAC, researchers can write a reproducible, dynamically generated draft of their intended manuscript at the preregistration stage.
This already includes most of the typical sections, such as introduction, methods, and results.
These results are initially based on simulated data with the same structure as the data the authors expect to obtain from their experiments.
For guidance on how to simulate data, see @morrisUsingSimulationStudies2019, @paxtonMonteCarloExperiments2001, and @skrondalDesignAnalysisMonte2000, as well as the R packages [`simstudy`](https://cran.r-project.org/web/packages/simstudy/vignettes/simstudy.html), [@simstudy] and [`psych`](https://personality-project.org/r/psych/help/sim.html), [@psych].

Once the preregistration is submitted and real data have been collected or made available,
the document can be reproduced with a single command, thus updating the Results section to the final version.
Reproducibility is of utmost importance at this stage since the preregistration must produce valid results at two points in time, once before data collection and once after data collection.
As outlined before, reproducibility builds upon four pillars (version control, dynamic document generation, dependency tracking, and software management).<!-- CJ I don't understand why you repeat this here. Remove?-->
To use PAC the dangers to reproducibility we described must be  eliminated.<!-- CJ I don't understand why you repeat this here. Remove?-->

The idea of submitting code as part of a preregistration is not new [e.g. @wichertsDegreesFreedomPlanning2016].<!--CJ: Please keep this comment in the text. If possible, I would like to reference the upcoming many analysts paper here, for which I used PAC in the final revision.-->
A prominent preregistration platform, [The Open Science Framework](https://osf.io/), suggests submitting scripts alongside the preregistration of methods.
In an informal literature search (we skimmed the first 300 results of Google Scholar with the keywords `("pre registration"|"pre-registration"|preregistration)&(code|script|matlab|python|"R")`) we only found close to a dozen published articles that did include some form of script as part of their preregistration.
Though the notion of preregistering code has been around for a while [cf. @steegenMeasuringCrowdAgain2014],
it has not gained much traction---perhaps because, to date, this has constituted an extra non-standard step in the research process.
This tutorial integrates the preregistration of code into the reproducible research workflow
by encouraging researchers to preregister the whole manuscript as a dynamic document.

## Advantages of PAC Over Traditional Preregistration

We believe that pairing PAC with the workflow presented above offers several advantages over classical preregistration.
First, PAC is merely an intermediate stage of the final manuscript, thus sparing authors from writing, and editors and reviewers from evaluating, two separate documents.
Relatedly, writing the preregistration in the form of a research article has the advantage that researchers are usually familiar with this format.
By contrast, a preregistration template is a novelty for many.
Second, PAC is a tool for study planning.
A study can be carried out more efficiently if all steps are documented clearly than when every step is planned ad hoc.
Third, PAC removes ambiguity regarding the translation of verbal analysis plans into code.
PAC is more comprehensive by design because its completeness can be empirically checked with simulated data.
Evaluating the intended analysis code on simulated data will help identify missing steps or ambiguous decisions.
PAC therefore minimizes undisclosed researchers degrees of freedom more effectively than standard preregistration does [@bakkerEnsuringQualitySpecificity2020; @wichertsDegreesFreedomPlanning2016].
Fourth, despite its rigor, PAC accommodates partly exploratory data-dependent decisions.
Some decisions can be anticipated and formulated as conditional code.
Researchers can, for example, formulate conditions (e.g., in the form of if-else-blocks) under which they prefer one analysis type over the other.
For example, if distributional assumptions are not met, the code may branch out to employ robust methods; or, an analysis may perform automated variable selection mechanisms before running the final model.
Some necessary deviations from the preregistration only become apparent after seeing the data.<!--CJ I merged point 5 here; it's part of point 4-->
Version control makes it possible to commit specific changes to the planned analyses.
It is recommended to documented all such changes by maintaining a changelog, and referencing specific commits.<!--CJ I do this myself, I think it's a good idea.-->
Another example of data-dependent decisions are more explorative analyses, i.e. explorative factor analysis or machine learning.
Decisions that do not lend themselves to formulation in code, e.g. visual inspection, must still be described verbally or be treated as noted in the next section.
With PAC researchers can precisely calibrate how much they want to explore.<!--CJ This is vague-->
When communicating a finding as explorative, it may still make a difference if tens or hundreds of variables were considered.<!-- CJ Not clear what this refers to; you haven't given an example or anything! Cut?-->

## Deviating from the Preregistration and Exploration

We would like to note that PAC allows explicit comparison of the preregistration and the final publication.
Authors should retrospectively summarize and justify any changes made to the preregistered plan, e.g. in the discussion of the final manuscript [^summarize-changes].
During the analysis process, authors can additionally maintain a running changelog to explain changes in detail as they arise.
Each entry in the changelog should explain the reasoning behind the changes and link to the commit id that applied the changes.
This enables readers and reviewers to inspects individual changes and make an informed judgment about their validity and implications.

Deviations from the preregistration are sometimes maligned, as if encountering unexpected challenges invalidates a carefully crafted study [@szollosi_is_2020].
However, we share the common view that deviation from a preregistration is not a problem [@nosekPreregistrationHardWorthwhile2019];
rather, a faillure to disclose such deviations is a problem.
In fact, it is expected that most PACs will require some modification after empirical data becomes available.
Often, deviations provide an opportunity to learn from the unexpected.
Transparency is essential in this regard <!--CJ: it is not clear how the following argument justifies the need for transparency. Please fix--> because an empirical test of a theory can not test the theory alone but tests it in conjunction with auxiliary assumptions [@meehlTheoreticalRisksTabular1978].

For example, imagine that authors preregistered their intention to include both "working memory" and "fluid intelligence" as covariates in an experimental study, examining the effect of task novelty on reaction time.
When evaluating the planned analyses on the empirical data, these two covariates reveal high collinearity, thus biasing estimates and compromising statistical inference.
The authors decide to use PCA to extract common variance related to "intelligence", and include this component as a covariate instead.
This change pertains to an auxiliary assumption (that working memory and fluid intelligence are distinct constructs), but does not undermine the core theory (that task novelty affects reaction time).
Now imagine that a different researcher is interested in the structure of intelligence.
This change to the preregistration directly relates their theory of intelligence.
That researcher might thus interpret the same result as an explorative finding, suggesting that these aspects of intelligence are unidimensional.
<!--CJ not sure what this final line intended to argue. I removed it. Feel free to add a more satisfactory conclusion to this paragraph-->

[^summarize-changes]: In section [Preregistration as Code --- a Tutorial] we conducted an actual PAC and summarize the changes we make to the preregistered code in the discussion.

Another common misunderstanding is that preregistration, including PAC, precludes exploratory analyses.
We differentiate between two kinds of exploration, neither of which is limited by PAC.
The first, more traditional kind of exploration involves ad hoc statistical decisions and post hoc explanations of the results.
Such traditional exploratory findings should be explicitly declared in the manuscript to distinguish them from confirmatory findings [@NosekRevolution2018; @nosekPreregistrationHardWorthwhile2019].
The second kind involves rigorous exploration using methods that are standard in machine learning [@brandmaier22mlsem].
These models often involve dozens, if not hundreds of predictors, which makes it difficult to describe them verbally.
With PAC, such models can be preregistered clearly and in comprehensive detail.
We specifically recommend PAC for such exploratory statistical models.
The merit of preregistration in these cases is to communicate precisely how much exploration was done;
a piece of information that is crucial to assess e.g., whether the results might be overfit <!--CJ reference Hastie & Tibshirani-->.

## How to Write a Good PAC

When planning a study and writing a PAC, researchers ought to strive to meet three goals:<!--CJ: None of these goals relate explicitly to code. Is that a problem? Also: Shouldn't you mention the goals when you first introduce PAC?-->
First, the plan should be comprehensive.
That is, researchers need to state the research question and describe the study design and analysis in detail.
Second, the plan should be effective, meaning that the researchers should reasonably expect it to answer the research question at hand.
Third, the plan should be efficient; i.e., it should only require the necessary amount of resources.

### Comprehensiveness

Standards for scientific writing are designed to result in understandable and comprehensive research articles and can serve the same purpose for writing a preregistration.
For example, the APA standards [@apa7] require research articles describe a study in such detail that any researcher that is familiar with the field is able to repeat it.
These well-known standards thus provide a good starting point for the structure of a PAC.<!--CJ What are you arguing here? You've just convinced the reader that existing templates are not comprehensive enough. Now you're arguing that APA7 can be considered a preregistration template, and that it is comprehensive? I am a little confused here--> 
One particular advantage is introduction's and theoretical background's embedding in existing literature where that provide the basis for the study's design.<!--CJ This sentence has been missed by the language editor-->
Further inspiration may be taken from preregistration templates relevant to the type of study at hand,
because such templates address methodological rigor more comprehensively.
One factor that limits the comprehensiveness of published scientific reports is the word count imposed by journals.
If this is a concern, we suggest relegating the PAC to a reproducible appendix to keep the core manuscript concise.

Simulating data offers further benefits in ensuring the comprehensiveness of a preregistration.
It forces the researcher to consider conceptual data-analytic challenges before obtaining empirical data [@axelrodAdvancingArtSimulation1997].
It also allows researchers to verify that planned analyses are feasible and likely to run without failure [@braiekTestingMachineLearning2020] (e.g., does the code also work with missing values, extreme outliers, etc.).
Many errors and omissions are likely to come to light when evaluating the intended analysis code on simulated data.

One final method to ensure comprehensivenes of the PAC is peer review.
The value of PAC would be further enhanced by submitting it as a registered report,
a format in which articles are accepted for publication "in principle" based on peer review of the preregistered study plans [@nosekRegisteredReports2014; @chambersWhatNextRegistered2019].
In such registered report peer-reviewing a manuscript-styled PAC may well make it easier for the reviewers that ascertain and evaluate a comprehensive study plan.
Moreover, reviewers can suggest specific improvements to the planned analysis code.

### Effectiveness

If the comprehensiveness criterion is met, researchers can further check whether the planned analyses are effective.
Effectiveness can be defined in terms of ability to recover true population parameters, or power to detect a population effect.<!--CJ You have not said anything about a simulation yet, so this is kind of surprising-->.
Effectiveness can be assessed by means of a simulation study.
Once code has been written to simulate synthetic data, it is often a minor step to perform a simulation study.
It is sufficient to repeat the planned analyses many times, and plot the distribution of estimated parameters, or tabulate the proportion of replications in which a true effect was detected (power analysis).
The R command `replicate()` can be used to this end.
If an analysis fails to consistently recover the population parameters under simulated (idealized) conditions,
it would be unreasonable to expect this analysis to be effective under realistic conditions.
Because the simulation describes the researchers' expectations about real-world conditions,
it is the best way to assess the effectiveness of the analysis prior to collecting empirical data.
Simulating data requires researchers to explicate their assumptions about the data-generating mechanism in concrete computational terms; e.g. assumptions about the distributions of variables and how they relate to one another.
If a researcher experiences difficulty in specifying such assumptions, this may indicate that further literature research or exploratory pilot studies are required.

### Efficiency<!-- CJ I do not understand how 'efficiency' is different from effectiveness above. You cannot assess effectiveness by means of a single iteration. Moreover, efficiency has a specific statistical definition (pertaining to the amount of information required to estimate a parameter). Using this term can be highly confusing to our (probably statistically inclined) readership.-->

A single simulated dataset already helps to ensure that the plan is viable under certain circumstances.
However, repeated simulation with varying parameters, a so-called Monte Carlo simulation, lets researchers play through several scenarios.
Monte Carlo simulation lets researchers navigate several trade-offs regarding their resources.
In the simplest case, they can assess the power (the probability to detect an effect of a given size) for varying numbers of observations in the simulation to determine the required sample size for their study [@brandmaier2018precision].
@bakkerRecommendationsPreregistrationsInternal2020 found that a formal power analysis may help researchers to invest resources more efficiently.
Researchers may also evaluate how the data analysis copes with expected data complications, e.g. non-normal data [@harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].

### Planned Analyses as Functions

Although researchers may use any form to preregister their planned analyses (e.g., scripts),
we suggest writing three functions for each planned hypothesis: one to conduct the planned analysis, one to simulate the expected data, and to report the results.
Using functions makes the analysis more portable (i.e., it can easily be used for other datasets), and facilitates repeated evaluation, as is the case in a simulation study. 
The functions shown here do not contain executable code, but the interested reader can find working functions in the [online supplementary materials](https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation_funs.R) that power the example below.

It is difficult to write analysis code when it is not clear what the expected data will look like.
We therefore recommend first simulating a dataset that resembles the expected structure of the empirical data that will be used for the final analysis.
Dedicated packages to simulate data for specific analyses exist <!--CJ Reference here-->.
For linear models, simulating data is extremely simple:

```{r}
simulate_data <- function(n, effect_size){
  # Draw n samples from a normal distribution for predictor X
  x <- rnorm(n)
  # Calculate dependent variable Y as a function of population effect size and residual error
  y <- effect_size * x + rnorm(n)
  # Return a data.frame
  data.frame(x = x, y = y)
}
simulate_data(5, .2)
```

The general format of a simulation function might be as follows:

```{r, eval=FALSE}
simulate_data <- function(n, effect_size){
  # 1. warn users that the results are "fake"
  # 2. draw `n` samples with `effect_size`
  # 3. format and return in expected data format
}
```

Next, write a function to conduct the planned analysis.
This function should receive the data and compute all relevant results from it.
In the simplest case, an analysis function might already exist in R.
For the linear model above, the analysis function might be:

```{r}
lm(y ~ x, data = simulate_data(5, .2))
```

The general format of an analysis function might be:

```{r}
planned_analyis <- function(data){
  # 1. preprocess e.g. with `rowMeans(data)`
  # 2. conduct analysis e.g. with `t.test()`
  # 3. `return(results)`
}
```

As soon as we have written `planned_anaylsis()` and `simulate_data()` we can iteratively improve both functions, e.g. until `planned_analysis()` runs without error and recovers the correct parameters from `simulate_data()`.
The goal is to ensure that the output of `simulate_data()` works as input to the function `planned_analysis()`.
Then, using the simulated dataset as a placeholder, one can write a scientific report including all analysis code blocks and the presentation of results.

### Alternatives to simulated data

If simulating data proves to be too challenging, researchers can delay their preregistration until they have gathered data and repurpose the empirical data.
In doing so, the researchers must walk a fine but important line.
Of course, they must not analyze the data before they write the preregistration.
Nevertheless, they may obfuscate the real data (before actually looking at them) by shuffling the order of rows for each variable (independently of each other) in cross-sectional data or otherwise by creating synthetic data that allow them to test the function `planned_analysis()` without actually revealing the observed research outcomes.
Preregistration after data collection is common for secondary data analysis of data obtained by other research groups [@westonRecommendationsIncreasingTransparency2019] but not so much within the same research project.
We argue that it is still an eligible preregistration.
In fact, guidelines for clinical trials also recommend analysis of blinded data  to test the practicality of the preregistration [@ICH1998].
Since the data is obfuscated, virtually no information is revealed to the researchers, and they can not employ their degrees of freedom effectively to influence the results (see section above).
Shuffling leaves only univariate information, so this procedure is only applicable for bi- or multivariate hypotheses.
The researcher can still access the information about means or proportions (e.g., the number of participants belonging to group "A" are in the dataset), but remain uninformed about relations between variables (e.g., members of group "A" have a greater mean in variable "Z").
To be clear, any exploration of the data before preregistration (PAC or otherwise) renders the preregistration invalid and amounts to (albeit sometimes unintentional) scientific misconduct.
The aim is transparency and the line that shall not be crossed is learning about the research outcomes, but what this means differs for each research project.

Another approach that can replace but also complement simulated data is to conduct a pilot study [@thabaneTutorialPilotStudies2010] and use the collected data to preregister.
A pilot study has obvious advantages for study planning, since it lets the researcher evaluate the feasibility of many assumptions.
However, we must warn our readers, that while piloting is more traditional than our approach of obfuscating the data before preregistration, it is not innocuous to use the data resulting from the pilot study.

When the researchers are satisfied with the function `planned_analysis()`, they can think about the way the would they would like to report the analysis results via tables, plots, and text.
The implementation of this reporting should be in the function `report_analysis()`.

```{r, eval=FALSE}
report_analysis <- function(results){
  # 1. create markdown tables from results
  # 2. conditionally interpret results e.g. if(p < .025)"Result is significant."
  # (optional) visualize results
  # 3. return results section formatted in markdown
}
```

This function should again accept the output of `planned_analysis()` as input.
The output of this function should be formatted in Markdown.
The idea is to automatically generate the full results section from the analysis.
This way, the preregistration not only specifies the computation but also how the its results are reported.
Various packages automatically generate well-formatted Markdown outputs of statistical reports or even entire tables of estimates or figures directly from R goal to help with this objective.
Packages like [`pander`](https://github.com/Rapporter/pander) [@pander], [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf) [@stargazer], [`apaTables`](https://dstanley4.github.io/apaTables/articles/apaTables.html) [@apaTables] and [`papaja`](http://frederikaust.com/papaja_man/) [@papaja] help you to create dynamically generated professional looking results.
The package [`report`](https://github.com/easystats/report) [@report] is particularly noteworthy because it not only generates tables but also a straightforward interpretation of the effects as actual prose (e.g., it verbally quantifies the size of an effect).

Ideally, these three functions can be composed to create a "fake" results section, e.g. when composed to `report_analysis(planned_analysis(simulate_data()))` or `simulate_data() %>% planned_analysis() %>% report_analysis()` outputs a results section.

To summarize, researchers need to write three functions, `planned_analysis()`, `simulate_data()`, and `report_analysis()` and embed these into a manuscript that serves as a preregistration.
After they gathered the actual data, they can replace the simulated data, render the dynamic manuscript (therefore run `planned_analysis()` on the actual data), and write the discussion.

#### When Is PAC Applicable?

Every study that can be preregistered and ultimately uses computer code for the statistical analysis, can also be preregistered with code.
The only difference is that PAC demands a comprehensive and complete analysis plan.
With traditional preregistration, researchers may (unintentionally) omit details and thus introduce decisions that appear to be made after having seen the data.
We emphasize again that deviations from the preregistration are not only possible but likely.
The advantage of PAC is transparency.
All deviations from the planned analysis as implemented in code are tracked, and the final analysis can be easily compared to the PAC.
Importantly, we see PAC as a means to increase transparency not as a means to limit researchers in their statistical analyses.
In practice, researchers may not feel ready to decide on every detail of the analysis plan before data collection.
In such cases, we recommend making decisions based on the best available knowledge, and if these decisions still prove infeasible, make simple assumptions and revise and explain them later as described in Section [Deviating from the preregistration and exploration] in the final manuscript.

Beyond standard preregistrations in psychological research, we believe that two applications specifically lend themselves to PAC.
First, pregistrations of clinical trials (called statistical analysis plans, @ICH1998) typically describe analyses in exhaustive detail and typically contain a detailed description of how results will be presented, including shells of tables and graphics [@yuanGuideStatisticalAnalysis2019].
PAC may significantly reduce the required workload while maintaining (and exceeding) the required standards for preregistering a clinical trial.

Second, preregistering exploratory statistical models (i.e., those with large numbers of competing models or those inspired by machine learning) is hardly feasible with standard preregistrations since they are too complex to describe and depend strongly on their software implementation.
PAC, however, captures the precise algorithmic model, including its software implementation and is ideal for preregistering these models [@brandmaier22mlsem].

# Preregistration as Code: Tutorial

We ahve argued that PAC has several advantages over classic preregistration and have outlined its implementation.
To illustrate how PAC works in practice and help researchers to implement PAC themselves, we have conducted one that serves as a guiding example.
We will use an exemplary research question that was based on openly available data and merely chosen for simplicity of illustration:

> "Is there a mean difference in the personality trait 'Machiavellism' between self-identified females and males?"

Again, we propose a preregistration format that closely resembles a classic journal article but uses simulated data and dynamic document generation to create a document that starts out as a preregistration and eventually becomes the final report.
The complete preregistration source is available in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/blob/main/preregistration.Rmd).
In this section, we show code excerpts of this preregistration (formatted in monospace) and explain the rationale behind them.

```{r, echo=FALSE}
show <- function(x, wrap = TRUE, cat = TRUE, fence = TRUE){
  stopifnot(length(x) == 1L)
  if(wrap)x <- str_c(str_wrap(str_split(x, "\\n\\n")[[1]]), collapse = "\n\n")
  if(!isFALSE(fence)){
    if(isTRUE(fence))fence <- "markdown"
    fence <- str_c("````", fence)
    x <-
      str_c(fence,
            x,
            "````",
            sep = "\n",
            collapse = "\n")
  }
  if(cat){
    cat(x)
    return(invisible(x))
  } else {
    return(x)
  }
}
show_prereg <-
  function(section,
           file = "preregistration.Rmd",
           wrap = TRUE,
           cat = TRUE,
           fence = TRUE) {
    prereg <- readr::read_lines(here::here(file))
    start <- stringr::str_c("<!--begin-", section, "-->")
    end <- stringr::str_c("<!--end-", section, "-->")
    start_pos <- which(prereg == start)
    end_pos <- which(prereg == end)
    if (length(start_pos) != 1L || length(end_pos) != 1L)
      stop("No section '", section, "' found.")
    excerpt <-
      str_c(prereg[(start_pos + 1):(end_pos - 1)], collapse = "\n")
    show(excerpt, wrap = wrap, cat = cat, fence = fence)
  }
```

As usual, the authors state why they are interested in their research question in the "Introduction" section and provide the necessary background information and literature to understand the context and purpose of the research question.
This example is drastically shortened for illustration purposes:

```{r, echo=FALSE, results='asis'}
show_prereg("theory")
```

After researchers have provided the research question, they typically proceed to explain how they want to study it.
For simplicity, we will use already published data that we have not yet analyzed:

```{r, echo=FALSE, results='asis'}
show_prereg("method")
```

We choose the following statistical procedure because many researchers are familiar with it[^caution-mww]:

```{r, echo=FALSE, results='asis'}
show_prereg("ttest")
```

[^caution-mww]: The t-test and Mann-Whitney-Wilcoxon test are arguably the most often used hypothesis tests (according to [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005] reports that 26% of all studies employed a t-test and 27% employed a rank-based alternative in the *New England Journal of Medicine* in 2005).
The analytical strategy presented here is, in fact, suboptimal concerning several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in the presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]). The interested reader can use the [provided code for the simulation](https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R) to verify that the t-test provides unbiased effect sizes but the Mann-Whitney-Wilcoxon overestimates effect size with increasing sample size and skewness.

The methods section is the translation of the following `planned_analysis()` function:

```{r planned_analysis}
```

This function illustrates two advantages of PAC.
First, a PAC can easily include data-dependent decisions by creating different analysis branches under different conditions.
Second, it highlights how difficult it is to describe a statistical analysis precisely.
The same verbal descriptions may be implemented differently by different persons depending on their statistical and programming knowledge and their expectations and beliefs.
One example would be using the function `wilcox.test` instead of the combinations of the functions `rank` and `t.test`.
Either of them is a valid implementation of the Mann–Whitney–Wilcoxon test, but the first assumes equal variance.
In contrast, the second applies Welch's correction by default and hence is robust even with unequal variances across groups [@zimmermanRankTransformationsPower1993].
Mentioning every such minute implementation detail is almost impossible and would result in overly verbose preregistrations.
Still, these details can make a difference in the interpretation of statistical results and, thus, represent undisclosed researchers' degrees of freedom.

Together with the function `simulate_data()` (not shown here), the function `planned_analysis()` can be used to justify the planned sample size.
To that end, `simulate_data()` is repeatedly called with increased sample sizes and the proportion of significant results (power) is recorded.
The results for such a Monte Carlo simulation for this example are visualized in Figure \@ref(fig:power).
The code for this power analysis can be found in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R).
The next snippet shows how we integrated the results dynamically into the preregistration (the origin of the R-variables `minn`, `choosen_power`, and `choosen_d` is not shown).

```{r, echo=FALSE}
res <-
  read_csv(here::here("data", "simulation_results.csv"), col_types = "ddddddd")
choosen_power <- .8
choosen_d <- .2
minn <- filter(res, d == choosen_d, power > choosen_power) %>% 
  filter(n == min(n))
```

```{r, include=FALSE}
fig_caption <- str_c("Results of simulation for the power analysis. The cross indicates the sample size that archives ", choosen_power*100, "% assuming a Cohen's D of ", choosen_d, ".")
```

```{r power, echo=FALSE, fig.cap=fig_caption}
res %>%
  ggplot(aes(n, power, color = d, group = d)) +
  geom_line() +
  geom_point(
    data = minn,
    size = 3,
    color = "black",
    shape = 3
  ) +
  theme_minimal() +
  scale_color_viridis_c() +
  theme(legend.title = element_text()) +
  labs(y = "Power\n(Proportion of significant results)",
       x = "Sample size",
       color = "Cohen's D") +
  NULL
```

```{r, echo=FALSE, results='asis'}
show_prereg("power")
```

Monte Carlo simulations are, of course, not only applicable for this analysis method and also allow researchers to investigate further relevant properties of their analysis method beyond power [@brandmaier2018precision; @harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].

We implemented a mechanism that only uses simulated data when the actual data are not yet available (in this example, if the file `data/data.csv` does not exist) for the results section.
This mechanism also warns readers if these results are based on simulated data.
The warning is colored red to avoid any confusion between mock and actual results.
As soon as the actual data are available, the simulated data are no longer used, and the results represent the actual empirical results of the study.

```{r, echo=FALSE, results='asis'}
show_prereg("results", wrap = FALSE)
```

Naturally we did not access the data when we initially wrote this code.
We therefore did not know the exact format the data would have.
This means that we did need to change our preregistration after accessing the data to include i.e. the recoding of gender (lines 17-22) and the items (lines 23-26).[^prereg-changes]

[^prereg-changes]: We invite the reader to evaluate the changes we made to the preregistered code.
Either on [GitHub](https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg...main) → "Files changed" or directly in Git with `git diff v0.0.1.1-prereg preregistration.Rmd`.

The last thing we need to preregister is the reporting of our results with the combination of the functions `planned_analysis()` and `report_analysis()`.

```{r, echo=FALSE, results='asis'}
show_prereg("report", wrap = FALSE)
```

This is an example of how the results could be reported (based on simulated data):

```{r, eval=TRUE, echo=-1, results='hide'}
# this is a fake chunk, the faked output comes the chunk below
report_analysis(planned_analysis(simulate_data(900, 8, 0.3, 10)))
```

```{r, echo=FALSE, results='asis'}
set.seed(1234)
show(report_analysis(planned_analysis(simulate_data(900, 8, 0.3, 10)), cat = FALSE), fence = "")
```

This example of a preregistration covers a single study with a single hypothesis.
To organize studies with multiple hypotheses, we suggest multiple `planned_analysis()` and `report_analysis()` functions (possibly numbered in accordance with the hypotheses, e.g. 1.2, 2.3 etc.).
Computational steps that apply to multiple analyses should be extracted into their own function, for example, those that perform data cleaning or extract results from a statistical model in an identical way.
Preregistrations that cover multiple distinct data sources may employ multiple `simulate_data()` functions.
These are merely suggestions, and researchers are encouraged to find their own way of how to best organize their analysis code.

Of course, it is likely that the preregistered code has to be adapted to the actual data.
Changing the preregistration must be discussed.
This is our summary of what we changed:

```{r, echo=FALSE, results='asis'}
show_prereg("discuss-changes")
```

But readers can inspect and judge the changes for themselves on [GitHub](https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b).

The interested reader can find the example rendered as a [PDF file with real data](https://github.com/aaronpeikert/repro-tutorial/files/7309455/preregistration.pdf) in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/releases/tag/v0.0.3.1-results) and can view the changes we made since preregistering it on this [GitHub page](https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b).

After researchers are satisfied with their draft preregistration, they should create an archived and unchangeable version of the project that serves as the preregistration.
zenodo.org [@zenodo] is a publicly funded service provider that archives digital artefacts for research and provides digital object identifiers (DOI) for these archives.
While the service is independent of GitHub---in terms of storage facilities and financing---you can link GitHub and zenodo.org.
Please note that you can only link public GitHub repositories to zenodo.org.
You may log into zenodo.org through your GitHub account.
To log in with your GitHub account:

> Navigate to https://zenodo.org/login/ →  Log in with GitHub

\noindent{}To link zenodo.org and GitHub

> Log into zenodo.org → Account → GitHub[^zenodo-github]

\noindent{}Or:

> Navigate to https://zenodo.org/account/settings/github/

[^zenodo-github]: https://zenodo.org/account/settings/github/

After you have linked a GitHub repository, you trigger the archival by creating a GitHub release.
To create GitHub release, navigate to GitHub:

```{r, eval=FALSE}
usethis::browse_github()
```

Then click on Releases → Draft a new release.
Here you can add all relevant binary files but at least a rendered version of the manuscript and the Docker image.

Taken together, we realize that PACs are challenging.
Not every research question or project is suitable for preregistration in general, and many analytical strategies could prove impractical for a PAC.
Therefore, PAC should not be seen as an attempt to invalidate classical preregistration or non-preregistered research but as a rather complementary approach.
We are convinced that many reasons can justify changing the preregistered code, but equally certain that readers should judge for themselves weather they find these deviations from the preregistration meaningful.
PAC enables just that.
Researchers can change their plans and possibly defend their decision to do so, e.g., in the discussion, while readers of the manuscript can clearly see what was changed through version control.

# Discussion

Increased automation is increasingly recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019], and therefore this workflow fits nicely together with other innovations that employ automation, like machine-readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
Automated research projects promise a wide range of applications, among them PAC [potentially to be submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], direct replication [@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], executable research articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations such as the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].

Central to these innovations is a property we call "reusability," fully promoted by the present workflow.
Reusable code can run on different inputs from a similar context and produce valid outputs.
This property is based on reproducibility but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] such that it is _built-for-reuse_ [@al-badareenReusableSoftwareComponents2010].
The reproducible workflow we present here is heavily automated and hence promotes reusability.
Furthermore, adhering to principles of reusability typically removes errors in the code and thus increases the likelihood that the statistical analysis is correct.
Therefore reproducibility facilitates traditional good scientific practices and provides the foundation for promising innovations.

## Summary

This paper demonstrated how the R package `repro` supports researchers in creating reproducible research projects, including reproducible manuscripts.
These are important building blocks for transparent and cumulative science because they enable others to reproduce statistical and computational results and reports later in time and on different computers.
The workflow we present here rests on four software solutions, (1) version control, (2) dynamic document generation, (3) dependency tracking and (4) software management to guarantee reproducibility.
We first demonstrated how to create a reproducible research project.
Then, we illustrated how such a project could be reproduced---either by the original author and/or collaborators or by a third party.

We finally presented an example of how the rigorous and automated reproducibility workflow introduced by `repro` may enable other innovations, such as preregistration as code (PAC).
In PAC the entire reproducible manuscript, including planned analyses and results based on simulated data, is preregistered.
This way, every use of a researchers' degree of freedom is disclosed.
Once real data is gathered, the reproducible manuscript is (re-)created with the real data.
PAC only becomes possible because reproducibility is ensured and leverages version control and dynamic document generation as key features of the workflow.

## Limitations

We have to distinguish two kinds of limitations.
First, some limitations are by design and are unlikely to be changed soon.
Our workflow inherits these from the software it relies on and the fundamental design principles these share with the workflow and `repro`.
Second, there are other limitations that we hope development efforts of the open-source community and, concerning `repro`, we will solve

As a principle, this workflow is incompatible with steps that cannot be automated.
This principle is at odds with some ingrained habits of researchers to mix and match manual and automated steps in data analysis.
To allow for automation, many researchers will have to search for alternative software.  
The automation-friendly software we present here has several technical but critical limitations.
`Git` can track any files (files larger than 1 GB only with extensions) but can only compare text files (with endings like, `.txt`, `.csv`, `.R`, `.py`,  or `.Rmd`), not binary files (with endings like `.docx`, `.exe`, or `.zip`).
With `R Markdown`, tables and graphics are dynamically generated from code but cannot be crafted by hand.
`Make` can orchestrate any programmable software but fails to automate software that will only obey a user clicking on the right buttons.
`Docker` can ship software that runs on Linux and can be automatically installed, but that requirement precludes much commercial or closed-source software.

This move away from software that has served researchers well for decades is understandably difficult and presents us with a conundrum.
On the one hand, we firmly believe that automated reproducibility makes research more productive and collaboration easier.
But, on the other hand, we expect researchers to invest considerable time in learning new tools and to persuade their collaborators to do the same.
Three arguments reconcile this apparent paradox.
First, this change will not happen all at once.
Automated reproducibility is an ideal that we believe has many advantages, but it is not an all-or-nothing decision.
Researchers can pick up one skill at a time and then help their fellow collaborators to do the same.
Second, the upfront investment is required once (and efforts such as `repro` are underway to reduce it) and will pay dividends over many research projects.
Third, the move towards open software for research offers several benefits beyond enabling automated reproducibility [@schaffnerFutureScientificJournals1994; @fitzgeraldTransformationOpenSource2006; @chaldecottHistoryScientificTechnical1965; @sonnenburgNeedOpenSource2007].

`repro`'s primary goal is to increase the user-friendliness of reproducibility tools for researchers.
Despite our best efforts to ensure this goal, the package is still in development and undoubtedly has limitations.
The workflow we present here introduces several software solutions.
This sheer number of dependencies represents in itself a threat to long-term reproducibility.
To benefit from automatic and convenient reproduction, researchers must use `Git`, `Make`, and `Docker`.
However, `Git` and `Make` are themselves included in the `Docker` image created by `repro`.
Researchers can therefore employ the `Docker` image manually to download the `Git` repository and execute `Make` for full reproduction.
In other words, the only hard requirement for reproduction and therefore its Achilles' heel, is `Docker`.
The `Docker` approach has two vulnerabilities.
First, and more importantly, the `Docker` image for the project and the `Git` repository have to remain available.
The `Dockerfile` (the plain text description to build a `Docker` image), as opposed to the image, is insufficient because it relies on too many service providers (e.g., Microsoft R Application Network, Ubuntu Package Archive).
We recommend archiving the `Git` repository and the `Docker` image with zenodo.org, a non-profit long-term storage for scientific data.
The necessary steps for archival on zenodo.org are described at the end of Section [Preregistration as Code --- a Tutorial].
The second vulnerability is that even if the existence of the `Docker` image and `Git` repository is guaranteed, future researchers still require software to run the image.
To that end, they can either rely on `Docker` itself or `Docker`-compatible alternatives (e.g., `CoreOS rkt`, `Mesos Containerizer`, `Singularity`).
The only way to remove the reliance on such external software is to turn the `Docker` image into a full operating system that subsequently can be installed and run on almost any modern computer.
This process is technically possible and would guarantee reproducibility for decades without any software dependency, assuming hardware that conforms to the x86 instruction set architecture continues to be available.
However, this process requires much technical knowledge and is currently not facilitated by `repro`.
In any case and without such precautions, do the `R Markdown`, the `Makefile` and the `Dockerfile` provide information that allows researchers to trace the computational steps and recreate the computational environment manually.
The `Makefile`, for example, is written in a way that researchers can manually trace the dependencies and execute commands in the right order, in case they are unable to run `Make` for some reason.
Hypothetically, even if `Docker` were to become unavailable one day, the `Dockerfile` still serves as unambiguous documentation of how the original system was set up and may help future users to create a software environment that closely resembles the original.

## Outlook

Open science practices are a continually evolving field where technical innovations foster changes in research practice.
Open data are much more widespread thanks to online storage facilities; preregistration is possible because there are preregistration platforms and so forth.
Similarly, we hope that fully automatic reproduction, e.g., with `repro` as a technical innovation, increase scientific rigor and efficiency.

In practice, this ideal of a fully automatic reproduction of research projects can conflict with the wide range of demands for more user-friendly and powerful software.
Some may find `Make` too complicated or that `Docker` requires too much storage space.
Yet others may find that they require other programming languages or want to scale their computation across hundreds of computers, e.g., via high-performance computing clusters or cloud computing.

`repro` was designed modularly to meet many such demands.
At the moment, `repro` only supports the combination of `R Markdown`, `Git`, `Make`, and `Docker`.
However, there are alternatives for each of these elements that may fit better into an individual research project.
`R Markdown` could be complemented or replaced by a dynamic Microsoft Word document with the help of `officer` [@officer] or `officedown` [@officedown] to accommodate a wider range of journal submission standards.
Instead of using formal version control with `Git`, `repro` could automatically save snapshots for increasing user-friendliness.
`Make` could be replaced by the more R-centered alternative `targets` for more convenience.
`Docker` could be combined with `renv` [@R-renv] to control the package versions precisely (our approach fixes the date, `renv` the exact package version).
Alternatively, `Docker` could be replaced by the more lightweight `renv` if no dependencies outside of R are considered crucial.
`Docker` does not satisfy the requirements of many HPC environments, but `Singularity` was designed to avoid this limitation while still being compatible with `Docker` images.

`repro`'s modular structure allows such alternative workflows, though they have not yet been implemented.
Depending on the demand by users, we will implement some of them in `repro` and hope for broad adoption of computational reproducibility in the near future.

# References